---
title: "16S itag analyses Fujimoto et al., submitted to Journal of Great Lakes Research"
author: "Vincent Denef"
date: "April 25, 2016"
output:
  md_document:
    toc: true
    variant: markdown_github
---
    
```{r global_options, include=FALSE}

knitr::opts_chunk$set(
  fig.width = 7, 
  fig.height = 5, 
  fig.align = 'center', 
  fig.path = 'Figures/',
  warning = FALSE, 
  message = FALSE
)

```

### Load Libraries
```{r load_libs, include=FALSE}
library(phyloseq)
library(ggplot2)
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(DESeq2)
library(pegas)
library(pgirmess)
library(multcomp)
library(multcompView)
library(grid)
library(gridExtra)
theme_set(theme_bw())

source("/Users/vdenef/Documents/UMich/Digital_Notebooks/GitHub/MicrobeMiseq/R/miseqR.R")
source("/Users/vdenef/Documents/UMich/Digital_Notebooks/GitHub/LM13_DNA/LM13_DNA_function.R")

setwd("/Users/vdenef/Documents/UMich/Digital_Notebooks/GitHub/LM13_DNA/")

### Set our theme for our plots down the road
set_ggtheme <- theme_set(theme_bw() + 
                    theme(plot.title = element_text(face="bold", size = 20),  #Set the plot title
                    strip.text.x = element_text(size=12, face="bold"),  #Set the facet titles on x-axis 
                    strip.text.y = element_text(size=12, face="bold"),  #Set the facet titles on x-axis 
                    strip.background = element_blank(),  #Set the facet background to no background
                    axis.title.x = element_text(face="bold", size=16),  #Set the x-axis title
                    axis.title.y = element_text(face="bold", size=16),  #Set the y-axis title
                    axis.text.x = element_text(colour = "black", size=14),  #Set the x-axis labels
                    axis.text.y = element_text(colour = "black", size=14),  #Set the y-axis labels
                    legend.title = element_text(size=14, face="bold"),  #Set the legend title 
                    legend.text = element_text(size = 14),  #Set the legend text
                    legend.position="right"))  #Default the legend position to the right

###set taxonomy plotting colors
group.colors <- c(Acidobacteria = "grey26", Actinobacteria = "royalblue", Alphaproteobacteria = "steelblue4", Armatimonadetes = "red", Bacteroidetes = "darkred", "BD1-5" = "chartreuse", Betaproteobacteria = "steelblue4", Caldiserica = "black","Candidate_division_BRC1" = "violetred4", "Candidate_division_JS1" = "aquamarine1", "Candidate_division_OD1" = "#6DDE88", "Candidate_division_OP3" = "hotpink", "Candidate_division_OP8" = "goldenrod1", "Candidate_division_OP11" = "chocolate4", "Candidate_division_SR1" = "tan3", "Candidate_division_TM7" = "skyblue1", "Candidate_division_WS3" = "magenta", Chlamydiae="violet", Chlorobi="cyan2", Chloroflexi="darkgreen", Cyanobacteria = "chartreuse3", Deferribacteres = "slateblue3", "Deinococcus-Thermus" = "violetred", Dictyoglomi = "cornsilk4", Deltaproteobacteria = "steelblue4", Elusimicrobia = "yellow3", Epsilonproteobacteria = "steelblue4", Fibrobacteres = "darkred", Firmicutes = "blue4", FGL7S = "palevioletred1", Fusobacteria = "slateblue1", Gammaproteobacteria = "steelblue4", Gemmatimonadetes="black", GOUTA4 = "plum1", "Hyd24-12" = "sienna2", JTB23 = "seashell2", Nitrospirae = "yellow1", "NPL-UPA2"="#652926", OC31 = "mediumpurple4", Planctomycetes = "mediumorchid3", Proteobacteria = "steelblue4", "SHA-109" = "lightsalmon3", SM2F11 = "lightskyblue2", SPOTSOCT00m83 = "orangered", Spirochaetae = "gold3", Tenericutes="pink", Thermotogae = "chocolate1", TA06 = "lightslateblue",TA18 = "rosybrown3", TM6 = "olivedrab", unclassified = "grey", Verrucomicrobia = "purple4", "WCHB1-60" = "palegreen", Other = "darkgrey", Cytophagia = "darkred", Sphingobacteria = "darkred",Flavobacteria = "darkred")
```

### Figure 1, panel C: Muskegon Lake buoy data   
```{r load buoy data}
#load buoy data downloaded from http://www.gvsu.edu/wri/buoy/data-generate.htm?concentration=1&date=4/23/13-11/3/13&time=0:00,12:00&x=date&y=clcon001,odo001,tp001,clcon002,tp004,clcon003,odo003,tp005
buoydata <- read.table("Input_data/wri_buoy_data.tsv",sep="\t",header=T,row.names=1,as.is=T)

#modify into dataframe with date in first column then value of parameter in second, and third colum states which parameter was measured
#also, convert fahrenheit to celcius for temps at 2m (T2), 7m (T7) and 8m (T8)
buoydata$Date <- row.names(buoydata)

T2<-data.frame(buoydata$Date,buoydata$T2)
T2$parameter<-"Temp 2m"
colnames(T2)<-c("date","value","parameter")
T2$value<-(T2$value-32)/9*5

T7<-data.frame(buoydata$Date,buoydata$T7)
T7$parameter<-"Temp 7m"
colnames(T7)<-c("date","value","parameter")
T7$value<-(T7$value-32)/9*5

T9<-data.frame(buoydata$Date,buoydata$T9)
T9$parameter<-"Temp 9m"
colnames(T9)<-c("date","value","parameter")
T9$value<-(T9$value-32)/9*5

C2<-data.frame(buoydata$Date,buoydata$CHL2)
C2$parameter<-"Chla 2m"
colnames(C2)<-c("date","value","parameter")

C5<-data.frame(buoydata$Date,buoydata$CHL5)
C5$parameter<-"Chla 5m"
colnames(C5)<-c("date","value","parameter")

DO2<-data.frame(buoydata$Date,buoydata$DO2)
DO2$parameter<-"DO 2m"
colnames(DO2)<-c("date","value","parameter")

DO8<-data.frame(buoydata$Date,buoydata$DO8)
DO8$parameter<-"DO 8m"
colnames(DO8)<-c("date","value","parameter")

buoy.df<-rbind(T2,T7,T9,C2,C5,DO2,DO8)
buoy.T<-rbind(T2,T7,T9)
buoy.C<-rbind(C2,C5)
buoy.D<-rbind(DO2,DO8)

```

```{r plot data}
buoy_plot_T <- ggplot(buoy.T,aes_string(x = "date",y = "value",group = "parameter",color = "parameter")) +
  scale_x_discrete(breaks = c("05/15/2013","07/15/2013", "08/15/2013", "09/23/2013", "11/02/2013"),labels = c("May 15", "Jul 15", "Aug 15", "Sep 23", "Nov 02"),drop = FALSE) +
  geom_line(size = 0.4) + #geom_point(size = 0.8) +
  ggtitle("") + xlab("") +
  ylab("°C") + 
  scale_color_manual(name = "Temp", values = c("Temp 2m" = "gray0", "Temp 7m" = "gray30", "Temp 9m" = "gray60"), breaks = c("Temp 2m", "Temp 7m" , "Temp 9m" ), labels = c("2 m","7 m","9 m")) +
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=30, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=0),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="right", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); buoy_plot_T

buoy_plot_C <- ggplot(buoy.C,aes_string(x = "date",y = "value",group = "parameter",color = "parameter")) +
  scale_x_discrete(breaks = c("05/15/2013","07/15/2013", "08/15/2013", "09/23/2013", "11/02/2013"),labels = c("May 15", "Jul 15", "Aug 15", "Sep 23", "Nov 02"),drop = FALSE) +
  geom_line(size = 0.4) + #geom_point(size = 0.8) +
  ggtitle("") + xlab("") +
  ylab("ug/L") + 
  scale_color_manual(name = "Chla", values = c("Chla 2m" = "darkgreen", "Chla 5m" = "darkslategrey"), breaks = c("Chla 2m" , "Chla 5m"), labels = c("2 m","5 m")) +
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=30, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=0),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="right", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); buoy_plot_C

buoy_plot_D <- ggplot(buoy.D,aes_string(x = "date",y = "value",group = "parameter",color = "parameter")) +
  scale_x_discrete(breaks = c("05/15/2013","07/15/2013", "08/15/2013", "09/23/2013", "11/02/2013"),labels = c("May 15", "Jul 15", "Aug 15", "Sep 23", "Nov 02"),drop = FALSE) +
  geom_line(size = 0.4) + #geom_point(size = 0.8) +
  ggtitle("") + xlab("") +
  ylab("mg/L") + 
  scale_color_manual(name = "DO", values = c("DO 2m" = "darkgoldenrod4", "DO 8m" = "darkgoldenrod3"), breaks = c("DO 2m", "DO 8m" ), labels = c("2 m","8 m")) +
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=30, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="right", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); buoy_plot_D        
```

```{r final figure}
pdf(file="Figures/Rplot_buoy.pdf", width= 2.5, height=3, pointsize= 8, colormodel="cmyk", useDingbats = FALSE)
grid.arrange(buoy_plot_T,buoy_plot_C,buoy_plot_D,ncol=1)
dev.off()
```

### Import Sequence Data
```{r Import sequence Data, include=FALSE}
###input data were obtained fromn the jgi output produced by their iTagger pipeline, which performs the following steps
#(1) CONTAM FILTER: Filter one or more contaminants using Duk (e.g. PhiX control, sequencing library adapter dimers, human contaminants, etc.).  For paired reads, the entire pair is filtered if one end-read has a high-scoring hit.
#(2) PRIMER TRIM: PCR primers (of the conserved region) are trimmed away.  For paired-end reads, both forward and reverse primers must be found, otherwise the entire pair is filtered.
#(3) ITERATIVE PAIR MERGING: Reads are trimmed as a pair, removing the last base from whichever end has the highest expected error in a window 5bp wide.  Reads are trimmed from mean + 3 standard deviations to mean - 3 standard deviations in 0.5 standard deviation steps.  After each trimming step, pairs are merged into single sequences with either Flash or Pandaseq.  Pairs which are not merged continue to the next round of trimming.  Paired reads which are not combined are discarded.
#(4) EXPECTED-ERROR FILTER: Merged reads are filtered if they have an expected number of errors which exceeds the threshold.  The config file indicates the maximum number of expected errors per 100bp. Note that Flash and Pandaseq produce different quality scores in the overlap-assembled regions.

###as JGI used the wrong file to connect sample names to read files, we need to correct these using the following commands
#TO BE ADDED




### loading mothur output with FWDB+silva taxonomy and sample metadata. 
## create phyloseq object
sharedfile <- "Input_data/jgi.contigs.good.unique.good.filter.unique.precluster.pick.pick.an.unique_list.shared"
taxfile <- "Input_data/jgi.contigs.good.unique.good.filter.unique.precluster.pick.pick.an.unique_list.0.03.cons.taxonomy"
mothurdata <- import_mothur(mothur_shared_file = sharedfile ,mothur_constaxonomy_file = taxfile )
sampledata <- read.table("Input_data/LMSampleDescription.tsv",sep="\t",header=T,row.names=1,as.is=T)
SAMPLE = sample_data(sampledata)
physeq = merge_phyloseq(mothurdata, SAMPLE)

### We need to change the taxonomy names: when using the fwdb taxonomy we need to add different headers after removing the last column, which contains no information except for 1 Verrucomicrobia taxon
tax_table(physeq) <- tax_table(physeq)[,-8] # this removes the final column
tax_table(physeq) <- cbind(tax_table(physeq),row.names(tax_table(physeq)))
colnames(tax_table(physeq)) <- c("Kingdom","Phylum","Class","Order","Family","Genus","Tribe","Species")

### removing non-bacterial reads (was already done in mothur, but just to be safe after merging taxonomies)
physeq <- subset_taxa(physeq, Kingdom == "Bacteria")

# Load physeq from workspace
#load(file = ".RData")

```

``` {r merge duplicates}
# Merge samples on Duplicates column
physeq_dupmerge <- merge_samples(physeq, "Duplicates")

# Weird bug, set taxa to rows
if (!taxa_are_rows(physeq_dupmerge)) {
  otu_table(physeq_dupmerge) <- t(otu_table(physeq_dupmerge))
}

## Fix metadata of merged duplicates 
# Grab unique lines (first entry) from inland phyloseq object metadata
unique_idx <- !(duplicated(sample_data(physeq)$Duplicates))
duplicate_metadata <- sample_data(physeq)[unique_idx, ]
# Assign new metadata to phyloseq object
sample_data(physeq_dupmerge) <- duplicate_metadata
physeq_dupmerge <- prune_taxa(taxa_sums(physeq_dupmerge) > 0, physeq_dupmerge)
```

```{r scaling, include=FALSE}

#check number of reads in each sample, differences in count are in part due differet numbers of chlorophyl reads depending on sample
sampsums <- data.frame(sums = sample_sums(physeq_dupmerge))

ggplot(sampsums, aes(x = sums)) + geom_histogram(binwidth = 10000)

min(sampsums)
max(sampsums)

# Scales reads to smallest library size which is 70,480
physeq.scale <- scale_reads(physeq_dupmerge, min(sample_sums(physeq_dupmerge)))
physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
```

```{r Proteobacteria and Bacteroidetes phylum change to class}

### ADD THE PROTEOBACTERIA CLASSES TO THE PHYLA NAME FIELD IN PHYLOSEQ OBJECT TAXONOMY 
phy <- data.frame(tax_table(physeq.scale))
Phylum <- as.character(phy$Phylum)
Class <- as.character(phy$Class)

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Proteobacteria"){
    if (Class[i] == "unclassified"){
      Phylum[i] <- Phylum[i]       
    } else {
      Phylum[i] <- Class[i]
    }
  } 
}

phy$Phylum <- Phylum
t <- tax_table(as.matrix(phy))

tax_table(physeq.scale) <- t

#same for physeq_dupmerge, but also add class name to Bacteroidetes phylum field
phy <- data.frame(tax_table(physeq_dupmerge))
Phylum <- as.character(phy$Phylum)
Class <- as.character(phy$Class)

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Proteobacteria"){
    if (Class[i] == "unclassified"){
      Phylum[i] <- Phylum[i]       
    } else {
      Phylum[i] <- Class[i]
    }
  } 
}

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Bacteroidetes"){
    if (Class[i] == "unclassified"){
      Phylum[i] <- Phylum[i] 
    } else {
      Phylum[i] <- Class[i]
    }
  } 
}

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Sphingobacteriia"){
      Phylum[i] <- "Sphingobacteria" 
    }
  } 

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Flavobacteriia"){
      Phylum[i] <- "Flavobacteria" 
    }
  } 

phy$Phylum <- Phylum
t <- tax_table(as.matrix(phy))

tax_table(physeq_dupmerge) <- t
```

##Figure 2: alpha diversity
```{r alpha diversity metrics calculation}
# For alpha diversity we will RAREFY our phyloseq object.  Our raw-data phyloseq object was:
physeq_dupmerge  # Raw Merged Samples
min(sample_sums(physeq_dupmerge))

# Following code from Michelle Berry's Butterflygut website: http://rstudio-pubs-static.s3.amazonaws.com/25722_9fc9bdc48f0d4f13b05fa61baeda57a0.html#alpha-diversity
# Rarefy to 70480 reads with replacement 100 times to estimate species richness
# Initialize matrices to store richness and evenness estimates
#richness <- matrix(nrow = 47,ncol = 100)  # Store richness:  We have 47 samples 
#row.names(richness) <- sample_names(physeq_dupmerge)
#evenness <- matrix(nrow = 47,ncol = 100)  #Store evenness
#row.names(evenness) <- sample_names(physeq_dupmerge)

# We want to be reproducible - so let's set the seed.
set.seed(3)

# For 100 replications, rarefy the OTU table to 70480 reads and store the richness and evennes estimates in our 2 matrices we just created. Run this once, afterwards load saved data
#for (i in 1:100) {
# r <- rarefy_even_depth(physeq_dupmerge, sample.size = 70480, verbose = FALSE, replace = TRUE)
#  rich <- as.numeric(as.matrix(estimate_richness(r, measures="Observed")))
#  richness[,i] <- rich
#  even <- as.numeric(as.matrix(estimate_richness(r, measures="InvSimpson")))
#  evenness[,i] <- even
#}

#write.table(richness, "Input_data/ObservedRichness100", row.names = TRUE)
#write.table(evenness, "Input_data/InvSimpson100", row.names = TRUE)

richness <- read.table("Input_data/ObservedRichness100", header = TRUE)
evenness <- read.table("Input_data/InvSimpson100", header = TRUE)

# Create a new matrix to hold the means and standard deviations of all the richness estimates
rich_stats = matrix(nrow= nrow(richness), ncol=1)
rich_stats[,1] = apply(richness, 1, mean)
#rich_stats[,2] = apply(richness, 1, sd)
rich_stats = data.frame(row.names(richness), rich_stats)
colnames(rich_stats) = c("samples","mean")
rich_stats$Test <- "Observed Richness"
rich_stats$Duplicates <- as.character(rich_stats$samples)
rich_stats <- merge(rich_stats,data.frame(duplicate_metadata),by="Duplicates")
for(i in 1:length(rich_stats$Season)){
  rich_stats$Depth_Seas[i]<-paste(as.character(rich_stats$Depth_C[i]),
                                  as.character(rich_stats$Season[i]))}
for(i in 1:length(rich_stats$Season)){
  rich_stats$Stat_Seas[i]<-paste(as.character(rich_stats$Station[i]),
                                  as.character(rich_stats$Season[i]))}
for(i in 1:length(rich_stats$Season)){
  rich_stats$Frac_Seas[i]<-paste(as.character(rich_stats$Fraction[i]),
                                  as.character(rich_stats$Season[i]))}

# Create a new matrix to hold the means and standard deviations of the evenness estimates
even_stats = matrix(nrow = nrow(evenness), ncol = 1)
even_stats[,1] = apply(evenness, 1, mean)
#even_stats[,2] = apply(evenness, 1, sd)
even_stats = data.frame(row.names(evenness), even_stats)
colnames(even_stats) = c("samples","mean")
even_stats$Test <- "Inverse Simpson"
even_stats$Duplicates <- as.character(even_stats$samples)
even_stats <- merge(even_stats,data.frame(duplicate_metadata),by="Duplicates")

###  Add SIMPSON'S EVENNESS!
#simps_even = as.data.frame(matrix(nrow = nrow(evenness), ncol = 2))
simps_even <- data.frame(matrix(nrow = nrow(evenness), ncol = 3))
colnames(simps_even) = c("samples","mean", "Test")
simps_even$samples <- even_stats$samples
simps_even$mean <-  even_stats$mean/rich_stats$mean
simps_even$Test <- "Simpson's Evenness"
simps_even$Duplicates <- as.character(simps_even$samples)
simps_even <- merge(simps_even,data.frame(duplicate_metadata),by="Duplicates")
for(i in 1:length(simps_even$Season)){
  simps_even$Depth_Seas[i]<-paste(as.character(simps_even$Depth_C[i]),
                                  as.character(simps_even$Season[i]))}
for(i in 1:length(simps_even$Season)){
  simps_even$Stat_Seas[i]<-paste(as.character(simps_even$Station[i]),
                                  as.character(simps_even$Season[i]))}
for(i in 1:length(simps_even$Season)){
  simps_even$Frac_Seas[i]<-paste(as.character(simps_even$Fraction[i]),
                                  as.character(simps_even$Season[i]))}
```

```{r richness plots}
###RICHNESS PLOTS
#adjust default color of boxplot outliers so that it matches boxplot color
update_geom_defaults("point", list(colour = NULL))
## 1. richness differences based on Season
rich_Season_KW <- kruskal.test(rich_stats$mean ~ as.factor(rich_stats$Season)) 
print(rich_Season_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
rich_Season_KW_MC <- kruskalmc(rich_stats$mean ~ as.factor(rich_stats$Season))  ## Defaults to P < 0.05
print(rich_Season_KW_MC)

# Test to find the letters to represent significance in a plot
rich_Season_KW_test <- rich_Season_KW_MC$dif.com$difference # select logical vector
names(rich_Season_KW_test) <- row.names(rich_Season_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
rich_Season_KW_letters <- multcompLetters(rich_Season_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
rich_Season_KW_sigs_dataframe <-  data.frame(as.vector(names(rich_Season_KW_letters$Letters)), as.vector(rich_Season_KW_letters$Letters))
colnames(rich_Season_KW_sigs_dataframe) <- c("Season", "siglabel")
rich_Season_KW_try <- merge(rich_stats, rich_Season_KW_sigs_dataframe, by = "Season")

rich_Season_KW_try$Season <-factor(rich_Season_KW_try$Season,levels=c( "Sp", "Su", "Fa"))

rich_Season_KW_sigs <- rich_Season_KW_try %>%
  select(Season, Season, siglabel) %>%
  distinct()

# finally, the plot, after first setting the order of the factor levels
alpha_title <- paste("Richness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
rich_Season_KW_try$Fraction <- factor(rich_Season_KW_try$Fraction,levels=c("B","E"))
rich_Season_KW_try$Season <- factor(rich_Season_KW_try$Season,levels=c("Sp","Su","Fa"))
rich_Season_KW_try$Station <- factor(rich_Season_KW_try$Station,levels=c("MLB","MM15","MM110"))
rich_Season_KW_try$Depth_C <- factor(rich_Season_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Richness_Season_boxplot.pdf", width= 5, height=5, pointsize= 8)
rich_Season_boxplot <- ggplot(rich_Season_KW_try, aes(x = Season, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 0)  + ggtitle("Season") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = rich_Season_KW_sigs, aes(label = siglabel, x = Season, y = .95), size = 2) +
  scale_color_manual(name = "season", values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),
                     labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                     labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
    scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("Sp", "Su", "Fa"),
                   labels=c("Spring\n(n=14)","Summer\n(n=17)", "Fall\n(n=16)")) + 
  xlab("") + ylab("Observed Richness") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); rich_Season_boxplot #top, right, bottom, left
dev.off()

## 2. richness differences based on Fraction
rich_Fraction_KW <- kruskal.test(rich_stats$mean ~ as.factor(rich_stats$Fraction)) 
print(rich_Fraction_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
rich_Fraction_KW_MC <- kruskalmc(rich_stats$mean ~ as.factor(rich_stats$Fraction))  ## Defaults to P < 0.05
print(rich_Fraction_KW_MC)

# Test to find the letters to represent significance in a plot
rich_Fraction_KW_test <- rich_Fraction_KW_MC$dif.com$difference # select logical vector
names(rich_Fraction_KW_test) <- row.names(rich_Fraction_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
rich_Fraction_KW_letters <- multcompLetters(rich_Fraction_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
rich_Fraction_KW_sigs_dataframe <-  data.frame(as.vector(names(rich_Fraction_KW_letters$Letters)), as.vector(rich_Fraction_KW_letters$Letters))
colnames(rich_Fraction_KW_sigs_dataframe) <- c("Fraction", "siglabel")
rich_Fraction_KW_try <- merge(rich_stats, rich_Fraction_KW_sigs_dataframe, by = "Fraction")

rich_Fraction_KW_try$Fraction <-factor(rich_Fraction_KW_try$Fraction,levels=c( "B", "E"))

rich_Fraction_KW_sigs <- rich_Fraction_KW_try %>%
  select(Fraction, Fraction, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Richness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
rich_Fraction_KW_try$Fraction <- factor(rich_Fraction_KW_try$Fraction,levels=c("B","E"))
rich_Fraction_KW_try$Season <- factor(rich_Fraction_KW_try$Season,levels=c("Sp","Su","Fa"))
rich_Fraction_KW_try$Station <- factor(rich_Fraction_KW_try$Station,levels=c("MLB","MM15","MM110"))
rich_Fraction_KW_try$Depth_C <- factor(rich_Fraction_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Richness_Fraction_boxplot.pdf", width= 5, height=5, pointsize= 8)
rich_Fraction_boxplot <- ggplot(rich_Fraction_KW_try, aes(x = Fraction, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 1)  + ggtitle("Fraction") +
  #geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = rich_Fraction_KW_sigs, aes(label = siglabel, x = Fraction, y = .95), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
 scale_x_discrete(breaks=c("B", "E"),
                   labels=c("FL\n(n=24)","PA\n(n=23)")) + 
   xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); rich_Fraction_boxplot #top, right, bottom, left
dev.off()


## 3. richness differences based on Station
rich_Station_KW <- kruskal.test(rich_stats$mean ~ as.factor(rich_stats$Station)) 
print(rich_Station_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
rich_Station_KW_MC <- kruskalmc(rich_stats$mean ~ as.factor(rich_stats$Station))  ## Defaults to P < 0.05
print(rich_Station_KW_MC)

# Test to find the letters to represent significance in a plot
rich_Station_KW_test <- rich_Station_KW_MC$dif.com$difference # select logical vector
names(rich_Station_KW_test) <- row.names(rich_Station_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
rich_Station_KW_letters <- multcompLetters(rich_Station_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
rich_Station_KW_sigs_dataframe <-  data.frame(as.vector(names(rich_Station_KW_letters$Letters)), as.vector(rich_Station_KW_letters$Letters))
colnames(rich_Station_KW_sigs_dataframe) <- c("Station", "siglabel")
rich_Station_KW_try <- merge(rich_stats, rich_Station_KW_sigs_dataframe, by = "Station")

rich_Station_KW_try$Station <-factor(rich_Station_KW_try$Station,levels=c( "MLB", "MM15", "MM110"))

rich_Station_KW_sigs <- rich_Station_KW_try %>%
  select(Station, Station, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Richness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
rich_Station_KW_try$Fraction <- factor(rich_Station_KW_try$Fraction,levels=c("B","E"))
rich_Station_KW_try$Season <- factor(rich_Station_KW_try$Season,levels=c("Sp","Su","Fa"))
rich_Station_KW_try$Station <- factor(rich_Station_KW_try$Station,levels=c("MLB","MM15","MM110"))
rich_Station_KW_try$Depth_C <- factor(rich_Station_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Richness_Station_boxplot.pdf", width= 5, height=5, pointsize= 8)
rich_Station_boxplot <- ggplot(rich_Station_KW_try, aes(x = Station, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 1)  + ggtitle("Station") +
  #geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = rich_Station_KW_sigs, aes(label = siglabel, x = Station, y = .95), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("MLB", "MM15", "MM110"),
                   labels=c("Estuary\n(n=10)","Near\n(n=18)", "Off\n(n=19)")) + 
  xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); rich_Station_boxplot #top, right, bottom, left
dev.off()


## 4. richness differences based on Depth_C
#rich_stats_off <- rich_stats[rich_stats$Station == "MM110" & rich_stats$Season == "Su", ]

rich_Depth_C_KW <- kruskal.test(rich_stats$mean ~ as.factor(rich_stats$Depth_C)) 
print(rich_Depth_C_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
rich_Depth_C_KW_MC <- kruskalmc(rich_stats$mean ~ as.factor(rich_stats$Depth_C))  ## Defaults to P < 0.05
print(rich_Depth_C_KW_MC)

# Test to find the letters to represent significance in a plot
rich_Depth_C_KW_test <- rich_Depth_C_KW_MC$dif.com$difference # select logical vector
names(rich_Depth_C_KW_test) <- row.names(rich_Depth_C_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
rich_Depth_C_KW_letters <- multcompLetters(rich_Depth_C_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
rich_Depth_C_KW_sigs_dataframe <-  data.frame(as.vector(names(rich_Depth_C_KW_letters$Letters)), as.vector(rich_Depth_C_KW_letters$Letters))
colnames(rich_Depth_C_KW_sigs_dataframe) <- c("Depth_C", "siglabel")
rich_Depth_C_KW_try <- merge(rich_stats, rich_Depth_C_KW_sigs_dataframe, by = "Depth_C")

rich_Depth_C_KW_try$Depth_C <-factor(rich_Depth_C_KW_try$Depth_C,levels=c( "S", "M", "D"))

rich_Depth_C_KW_sigs <- rich_Depth_C_KW_try %>%
  select(Depth_C, Depth_C, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Richness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
rich_Depth_C_KW_try$Fraction <- factor(rich_Depth_C_KW_try$Fraction,levels=c("B","E"))
rich_Depth_C_KW_try$Season <- factor(rich_Depth_C_KW_try$Season,levels=c("Sp","Su","Fa"))
rich_Depth_C_KW_try$Station <- factor(rich_Depth_C_KW_try$Station,levels=c("MLB","MM15","MM110"))
rich_Depth_C_KW_try$Depth_C <- factor(rich_Depth_C_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Richness_Depth_C_boxplot.pdf", width= 5, height=5, pointsize= 8)
rich_Depth_C_boxplot <- ggplot(rich_Depth_C_KW_try, aes(x = Depth_C, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 1)  + ggtitle("Depth") +
  #geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = rich_Depth_C_KW_sigs, aes(label = siglabel, x = Depth_C, y = .95), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("S", "M", "D"),
                   labels=c("Surface\n(n=30)","Chla max\n(n=2)","Deep\n(n=15)")) + 
  xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); rich_Depth_C_boxplot #top, right, bottom, left
dev.off()

```

```{r richness plot combination of factors to create multicomp letters based on Kruskall Wallis test between sub boxplots data}
###RICHNESS PLOTS
## 1. richness differences based on Season and Station combination
rich_Stat_Seas_KW <- kruskal.test(rich_stats$mean ~ as.factor(rich_stats$Stat_Seas)) 
print(rich_Stat_Seas_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
rich_Stat_Seas_KW_MC <- kruskalmc(rich_stats$mean ~ as.factor(rich_stats$Stat_Seas))  ## Defaults to P < 0.05
print(rich_Stat_Seas_KW_MC)

# Test to find the letters to represent significance in a plot
rich_Stat_Seas_KW_test <- rich_Stat_Seas_KW_MC$dif.com$difference # select logical vector
names(rich_Stat_Seas_KW_test) <- row.names(rich_Stat_Seas_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
rich_Stat_Seas_KW_letters <- multcompLetters(rich_Stat_Seas_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
rich_Stat_Seas_KW_sigs_dataframe <-  data.frame(as.vector(names(rich_Stat_Seas_KW_letters$Letters)), as.vector(rich_Stat_Seas_KW_letters$Letters))
colnames(rich_Stat_Seas_KW_sigs_dataframe) <- c("Stat_Seas", "siglabel")
rich_Stat_Seas_KW_try <- merge(rich_stats, rich_Stat_Seas_KW_sigs_dataframe, by = "Stat_Seas")

rich_Stat_Seas_KW_try$Stat_Seas <-factor(rich_Stat_Seas_KW_try$Stat_Seas,levels=c( "MLB Sp","MM15 Sp","MM110 Sp","MLB Su","MM15 Su","MM110 Su","MLB Fa","MM15 Fa","MM110 Fa"))

rich_Stat_Seas_KW_sigs <- rich_Stat_Seas_KW_try %>%
  select(Stat_Seas, Stat_Seas, siglabel) %>%
  distinct()

# finally, the plot, after first setting the order of the factor levels
alpha_title <- paste("Richness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
rich_Stat_Seas_KW_try$Fraction <- factor(rich_Stat_Seas_KW_try$Fraction,levels=c("B","E"))
rich_Stat_Seas_KW_try$Season <- factor(rich_Stat_Seas_KW_try$Season,levels=c("Sp","Su","Fa"))
rich_Stat_Seas_KW_try$Station <- factor(rich_Stat_Seas_KW_try$Station,levels=c("MLB","MM15","MM110"))
rich_Stat_Seas_KW_try$Depth_C <- factor(rich_Stat_Seas_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Richness_Seas_Stat_boxplot.pdf", width= 5, height=5, pointsize= 8)
rich_Seas_Stat_boxplot <- ggplot(rich_Stat_Seas_KW_try, aes(x = Stat_Seas, y = mean))  +
  geom_boxplot(outlier.size = 0)  + ggtitle("Season & Station") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = rich_Stat_Seas_KW_sigs, aes(label = siglabel, x = Stat_Seas, y = .95), size = 2) +
  scale_color_manual(name = "season", values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),
                     labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                     labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
    scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("MLB Sp", "MM15 Sp", "MM110 Sp","MLB Su", "MM15 Su", "MM110 Su","MLB Fa", "MM15 Fa", "MM110 Fa"),
                   labels=c("MLB Sp", "MM15 Sp", "MM110 Sp","MLB Su", "MM15 Su", "MM110 Su","MLB Fa", "MM15 Fa", "MM110 Fa")) + 
  xlab("") + ylab("Observed Richness") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); rich_Seas_Stat_boxplot #top, right, bottom, left
dev.off()

## 2. richness differences based on Fraction and Season
rich_Frac_Seas_KW <- kruskal.test(rich_stats$mean ~ as.factor(rich_stats$Frac_Seas)) 
print(rich_Frac_Seas_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
rich_Frac_Seas_KW_MC <- kruskalmc(rich_stats$mean ~ as.factor(rich_stats$Frac_Seas))  ## Defaults to P < 0.05
print(rich_Frac_Seas_KW_MC)

# Test to find the letters to represent significance in a plot
rich_Frac_Seas_KW_test <- rich_Frac_Seas_KW_MC$dif.com$difference # select logical vector
names(rich_Frac_Seas_KW_test) <- row.names(rich_Frac_Seas_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
rich_Frac_Seas_KW_letters <- multcompLetters(rich_Frac_Seas_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
rich_Frac_Seas_KW_sigs_dataframe <-  data.frame(as.vector(names(rich_Frac_Seas_KW_letters$Letters)), as.vector(rich_Frac_Seas_KW_letters$Letters))
colnames(rich_Frac_Seas_KW_sigs_dataframe) <- c("Frac_Seas", "siglabel")
rich_Frac_Seas_KW_try <- merge(rich_stats, rich_Frac_Seas_KW_sigs_dataframe, by = "Frac_Seas")

rich_Frac_Seas_KW_try$Frac_Seas <-factor(rich_Frac_Seas_KW_try$Frac_Seas,levels=c( "B Sp", "B Su", "B Fa", "E Sp", "E Su", "E Fa"))

rich_Frac_Seas_KW_sigs <- rich_Frac_Seas_KW_try %>%
  select(Frac_Seas, Frac_Seas, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Richness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
rich_Frac_Seas_KW_try$Fraction <- factor(rich_Frac_Seas_KW_try$Fraction,levels=c("B","E"))
rich_Frac_Seas_KW_try$Season <- factor(rich_Frac_Seas_KW_try$Season,levels=c("Sp","Su","Fa"))
rich_Frac_Seas_KW_try$Station <- factor(rich_Frac_Seas_KW_try$Station,levels=c("MLB","MM15","MM110"))
rich_Frac_Seas_KW_try$Depth_C <- factor(rich_Frac_Seas_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Richness_Frac_Seas_boxplot.pdf", width= 5, height=5, pointsize= 8)
rich_Frac_Seas_boxplot <- ggplot(rich_Frac_Seas_KW_try, aes(x = Frac_Seas, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 0)  + ggtitle("Fraction & Season") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = rich_Frac_Seas_KW_sigs, aes(label = siglabel, x = Frac_Seas, y = .95), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
 scale_x_discrete(breaks=c("B Sp", "B Su", "B Fa", "E Sp", "E Su", "E Fa"),
                   labels=c("FL Sp", "FL Su", "FL Fa", "PA Sp", "PA Su", "PA Fa")) + 
   xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); rich_Frac_Seas_boxplot #top, right, bottom, left
dev.off()


## 3. richness differences based on Station and Season
#  Resort the factors of Season & Fraction object to have primary sorting by station
rich_Stat_Seas_KW_try$Stat_Seas <-factor(rich_Stat_Seas_KW_try$Stat_Seas,levels=c( "MLB Sp","MLB Su","MLB Fa","MM15 Sp","MM15 Su","MM15 Fa","MM110 Sp","MM110 Su","MM110 Fa"))

rich_Stat_Seas_KW_sigs <- rich_Stat_Seas_KW_try %>%
  select(Stat_Seas, Stat_Seas, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Richness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
rich_Station_KW_try$Fraction <- factor(rich_Station_KW_try$Fraction,levels=c("B","E"))
rich_Station_KW_try$Season <- factor(rich_Station_KW_try$Season,levels=c("Sp","Su","Fa"))
rich_Station_KW_try$Station <- factor(rich_Station_KW_try$Station,levels=c("MLB","MM15","MM110"))
rich_Station_KW_try$Depth_C <- factor(rich_Station_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Richness_Stat_Seas_boxplot.pdf", width= 5, height=5, pointsize= 8)
rich_Stat_Seas_boxplot <- ggplot(rich_Stat_Seas_KW_try, aes(x = Stat_Seas, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 0)  + ggtitle("Station & Season") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = rich_Stat_Seas_KW_sigs, aes(label = siglabel, x = Stat_Seas, y = .95), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("MLB Sp", "MM15 Sp", "MM110 Sp","MLB Su", "MM15 Su", "MM110 Su","MLB Fa", "MM15 Fa", "MM110 Fa"),
                   labels=c("MLB Sp", "MM15 Sp", "MM110 Sp","MLB Su", "MM15 Su", "MM110 Su","MLB Fa", "MM15 Fa", "MM110 Fa")) +   xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); rich_Stat_Seas_boxplot #top, right, bottom, left
dev.off()


## 4. richness differences based on Depth_C and Season
#rich_stats_off <- rich_stats[rich_stats$Station == "MM110" & rich_stats$Season == "Su", ]

rich_Depth_Seas_KW <- kruskal.test(rich_stats$mean ~ as.factor(rich_stats$Depth_Seas)) 
print(rich_Depth_Seas_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
rich_Depth_Seas_KW_MC <- kruskalmc(rich_stats$mean ~ as.factor(rich_stats$Depth_Seas))  ## Defaults to P < 0.05
print(rich_Depth_Seas_KW_MC)

# Test to find the letters to represent significance in a plot
rich_Depth_Seas_KW_test <- rich_Depth_Seas_KW_MC$dif.com$difference # select logical vector
names(rich_Depth_Seas_KW_test) <- row.names(rich_Depth_Seas_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
rich_Depth_Seas_KW_letters <- multcompLetters(rich_Depth_Seas_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
rich_Depth_Seas_KW_sigs_dataframe <-  data.frame(as.vector(names(rich_Depth_Seas_KW_letters$Letters)), as.vector(rich_Depth_Seas_KW_letters$Letters))
colnames(rich_Depth_Seas_KW_sigs_dataframe) <- c("Depth_Seas", "siglabel")
rich_Depth_Seas_KW_try <- merge(rich_stats, rich_Depth_Seas_KW_sigs_dataframe, by = "Depth_Seas")

rich_Depth_Seas_KW_try$Depth_Seas <-factor(rich_Depth_Seas_KW_try$Depth_Seas,levels=c( "S Sp", "S Su", "S Fa", "M Su", "D Sp", "D Su", "D Fa"))

rich_Depth_Seas_KW_sigs <- rich_Depth_Seas_KW_try %>%
  select(Depth_Seas, Depth_Seas, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Richness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
rich_Depth_Seas_KW_try$Fraction <- factor(rich_Depth_Seas_KW_try$Fraction,levels=c("B","E"))
rich_Depth_Seas_KW_try$Season <- factor(rich_Depth_Seas_KW_try$Season,levels=c("Sp","Su","Fa"))
rich_Depth_Seas_KW_try$Station <- factor(rich_Depth_Seas_KW_try$Station,levels=c("MLB","MM15","MM110"))
rich_Depth_Seas_KW_try$Depth_C <- factor(rich_Depth_Seas_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Richness_Depth_Seas_boxplot.pdf", width= 5, height=5, pointsize= 8)
rich_Depth_Seas_boxplot <- ggplot(rich_Depth_Seas_KW_try, aes(x = Depth_Seas, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 0)  + ggtitle("Depth & Season") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = rich_Depth_Seas_KW_sigs, aes(label = siglabel, x = Depth_Seas, y = .95), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("S Sp", "S Su", "S Fa", "M Su", "D Sp", "D Su", "D Fa"),
                   labels=c("S Sp", "S Su", "S Fa", "M Su", "D Sp", "D Su", "D Fa")) + 
  xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); rich_Depth_Seas_boxplot #top, right, bottom, left
dev.off()
```


```{r evenness plots} 
###EVENNESS PLOTS
## 1. evenness differences based on Season
even_Season_KW <- kruskal.test(simps_even$mean ~ as.factor(simps_even$Season)) 
print(even_Season_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
even_Season_KW_MC <- kruskalmc(simps_even$mean ~ as.factor(simps_even$Season))  ## Defaults to P < 0.05
print(even_Season_KW_MC)

# Test to find the letters to represent significance in a plot
even_Season_KW_test <- even_Season_KW_MC$dif.com$difference # select logical vector
names(even_Season_KW_test) <- row.names(even_Season_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
even_Season_KW_letters <- multcompLetters(even_Season_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
even_Season_KW_sigs_dataframe <-  data.frame(as.vector(names(even_Season_KW_letters$Letters)), as.vector(even_Season_KW_letters$Letters))
colnames(even_Season_KW_sigs_dataframe) <- c("Season", "siglabel")
even_Season_KW_try <- merge(simps_even, even_Season_KW_sigs_dataframe, by = "Season")

even_Season_KW_try$Season <-factor(even_Season_KW_try$Season,levels=c( "Sp", "Su", "Fa"))

even_Season_KW_sigs <- even_Season_KW_try %>%
  select(Season, Season, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Evenness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
even_Season_KW_try$Fraction <- factor(even_Season_KW_try$Fraction,levels=c("B","E"))
even_Season_KW_try$Season <- factor(even_Season_KW_try$Season,levels=c("Sp","Su","Fa"))
even_Season_KW_try$Station <- factor(even_Season_KW_try$Station,levels=c("MLB","MM15","MM110"))
even_Season_KW_try$Depth_C <- factor(even_Season_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Evenness_Season_boxplot.pdf", width= 5, height=5, pointsize= 8)
even_Season_boxplot <- ggplot(even_Season_KW_try, aes(x = Season, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 0)  + #ggtitle(alpha_title) +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = even_Season_KW_sigs, aes(label = siglabel, x = Season, y = 0), size = 2) +
  scale_color_manual(name = "season", values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),
                     labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                     labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("Sp", "Su", "Fa"),
                   labels=c("Spring\n(n=14)","Summer\n(n=17)", "Fall\n(n=16)")) + 
  xlab("") + ylab("Simpson's Evenness") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); even_Season_boxplot #top, right, bottom, left
dev.off()


## 2. Evenness differences based on Fraction
even_Fraction_KW <- kruskal.test(simps_even$mean ~ as.factor(simps_even$Fraction)) 
print(even_Fraction_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
even_Fraction_KW_MC <- kruskalmc(simps_even$mean ~ as.factor(simps_even$Fraction))  ## Defaults to P < 0.05
print(even_Fraction_KW_MC)

# Test to find the letters to represent significance in a plot
even_Fraction_KW_test <- even_Fraction_KW_MC$dif.com$difference # select logical vector
names(even_Fraction_KW_test) <- row.names(even_Fraction_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
even_Fraction_KW_letters <- multcompLetters(even_Fraction_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
even_Fraction_KW_sigs_dataframe <-  data.frame(as.vector(names(even_Fraction_KW_letters$Letters)), as.vector(even_Fraction_KW_letters$Letters))
colnames(even_Fraction_KW_sigs_dataframe) <- c("Fraction", "siglabel")
even_Fraction_KW_try <- merge(simps_even, even_Fraction_KW_sigs_dataframe, by = "Fraction")

even_Fraction_KW_try$Fraction <-factor(even_Fraction_KW_try$Fraction,levels=c( "B", "E"))

even_Fraction_KW_sigs <- even_Fraction_KW_try %>%
  select(Fraction, Fraction, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Evenness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
even_Fraction_KW_try$Fraction <- factor(even_Fraction_KW_try$Fraction,levels=c("B","E"))
even_Fraction_KW_try$Season <- factor(even_Fraction_KW_try$Season,levels=c("Sp","Su","Fa"))
even_Fraction_KW_try$Station <- factor(even_Fraction_KW_try$Station,levels=c("MLB","MM15","MM110"))
even_Fraction_KW_try$Depth_C <- factor(even_Fraction_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Evenness_Fraction_boxplot.pdf", width= 5, height=5, pointsize= 8)
even_Fraction_boxplot <- ggplot(even_Fraction_KW_try, aes(x = Fraction, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 1)  + #ggtitle(alpha_title) +
  #geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = even_Fraction_KW_sigs, aes(label = siglabel, x = Fraction, y = 0), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
 scale_x_discrete(breaks=c("B", "E"),
                   labels=c("FL\n(n=24)","PA\n(n=23)")) + 
   xlab("") + ylab("") + 
  theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); even_Fraction_boxplot #top, right, bottom, left
dev.off()

## 3. Evenness differences based on Station
even_Station_KW <- kruskal.test(simps_even$mean ~ as.factor(simps_even$Station)) 
print(even_Station_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
even_Station_KW_MC <- kruskalmc(simps_even$mean ~ as.factor(simps_even$Station))  ## Defaults to P < 0.05
print(even_Station_KW_MC)

# Test to find the letters to represent significance in a plot
even_Station_KW_test <- even_Station_KW_MC$dif.com$difference # select logical vector
names(even_Station_KW_test) <- row.names(even_Station_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
even_Station_KW_letters <- multcompLetters(even_Station_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
even_Station_KW_sigs_dataframe <-  data.frame(as.vector(names(even_Station_KW_letters$Letters)), as.vector(even_Station_KW_letters$Letters))
colnames(even_Station_KW_sigs_dataframe) <- c("Station", "siglabel")
even_Station_KW_try <- merge(simps_even, even_Station_KW_sigs_dataframe, by = "Station")

even_Station_KW_try$Station <-factor(even_Station_KW_try$Station,levels=c( "MLB", "MM15", "MM110"))

even_Station_KW_sigs <- even_Station_KW_try %>%
  select(Station, Station, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Evenness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
even_Station_KW_try$Fraction <- factor(even_Station_KW_try$Fraction,levels=c("B","E"))
even_Station_KW_try$Season <- factor(even_Station_KW_try$Season,levels=c("Sp","Su","Fa"))
even_Station_KW_try$Station <- factor(even_Station_KW_try$Station,levels=c("MLB","MM15","MM110"))
even_Station_KW_try$Depth_C <- factor(even_Station_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Evenness_Station_boxplot.pdf", width= 5, height=5, pointsize= 8)
even_Station_boxplot <- ggplot(even_Station_KW_try, aes(x = Station, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 1)  + #ggtitle(alpha_title) +
  #geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = even_Station_KW_sigs, aes(label = siglabel, x = Station, y = 0), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +  scale_x_discrete(breaks=c("MLB", "MM15", "MM110"),
                   labels=c("Estuary\n(n=10)","Near\n(n=18)", "Off\n(n=19)")) + 
  xlab("") + ylab("") + 
  theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); even_Station_boxplot #top, right, bottom, left
dev.off()

## 4. Evenness differences based on Depth_C
#simps_even_off <- simps_even[simps_even$Station == "MM110" & simps_even$Season == "Su", ]

even_Depth_C_KW <- kruskal.test(simps_even$mean ~ as.factor(simps_even$Depth_C)) 
print(even_Depth_C_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
even_Depth_C_KW_MC <- kruskalmc(simps_even$mean ~ as.factor(simps_even$Depth_C))  ## Defaults to P < 0.05
print(even_Depth_C_KW_MC)

# Test to find the letters to represent significance in a plot
even_Depth_C_KW_test <- even_Depth_C_KW_MC$dif.com$difference # select logical vector
names(even_Depth_C_KW_test) <- row.names(even_Depth_C_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
even_Depth_C_KW_letters <- multcompLetters(even_Depth_C_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
even_Depth_C_KW_sigs_dataframe <-  data.frame(as.vector(names(even_Depth_C_KW_letters$Letters)), as.vector(even_Depth_C_KW_letters$Letters))
colnames(even_Depth_C_KW_sigs_dataframe) <- c("Depth_C", "siglabel")
even_Depth_C_KW_try <- merge(simps_even, even_Depth_C_KW_sigs_dataframe, by = "Depth_C")

even_Depth_C_KW_try$Depth_C <-factor(even_Depth_C_KW_try$Depth_C,levels=c( "S", "M", "D"))

even_Depth_C_KW_sigs <- even_Depth_C_KW_try %>%
  select(Depth_C, Depth_C, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("Evenness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
even_Depth_C_KW_try$Fraction <- factor(even_Depth_C_KW_try$Fraction,levels=c("B","E"))
even_Depth_C_KW_try$Season <- factor(even_Depth_C_KW_try$Season,levels=c("Sp","Su","Fa"))
even_Depth_C_KW_try$Station <- factor(even_Depth_C_KW_try$Station,levels=c("MLB","MM15","MM110"))
even_Depth_C_KW_try$Depth_C <- factor(even_Depth_C_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_Evenness_Depth_C_boxplot.pdf", width= 5, height=5, pointsize= 8)
even_Depth_C_boxplot <- ggplot(even_Depth_C_KW_try, aes(x = Depth_C, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 1)  + #ggtitle(alpha_title) +
  #geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = even_Depth_C_KW_sigs, aes(label = siglabel, x = Depth_C, y = 0), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("S", "M", "D"),
                   labels=c("Surface\n(n=30)","Chla max\n(n=2)","Deep\n(n=15)")) + 
  xlab("") + ylab("") + 
  theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); even_Depth_C_boxplot #top, right, bottom, left
dev.off()

```

```{r evenness plots combination of factors to create multicomp letters based on Kruskall Wallis test between sub boxplots data}
## 1. evenness differences based on Season and Station combination
even_Stat_Seas_KW <- kruskal.test(simps_even$mean ~ as.factor(simps_even$Stat_Seas)) 
print(even_Stat_Seas_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
even_Stat_Seas_KW_MC <- kruskalmc(simps_even$mean ~ as.factor(simps_even$Stat_Seas))  ## Defaults to P < 0.05
print(even_Stat_Seas_KW_MC)

# Test to find the letters to represent significance in a plot
even_Stat_Seas_KW_test <- even_Stat_Seas_KW_MC$dif.com$difference # select logical vector
names(even_Stat_Seas_KW_test) <- row.names(even_Stat_Seas_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
even_Stat_Seas_KW_letters <- multcompLetters(even_Stat_Seas_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
even_Stat_Seas_KW_sigs_dataframe <-  data.frame(as.vector(names(even_Stat_Seas_KW_letters$Letters)), as.vector(even_Stat_Seas_KW_letters$Letters))
colnames(even_Stat_Seas_KW_sigs_dataframe) <- c("Stat_Seas", "siglabel")
even_Stat_Seas_KW_try <- merge(simps_even, even_Stat_Seas_KW_sigs_dataframe, by = "Stat_Seas")

even_Stat_Seas_KW_try$Stat_Seas <-factor(even_Stat_Seas_KW_try$Stat_Seas,levels=c( "MLB Sp","MM15 Sp","MM110 Sp","MLB Su","MM15 Su","MM110 Su","MLB Fa","MM15 Fa","MM110 Fa"))

even_Stat_Seas_KW_sigs <- even_Stat_Seas_KW_try %>%
  select(Stat_Seas, Stat_Seas, siglabel) %>%
  distinct()

# finally, the plot, after first setting the order of the factor levels
alpha_title <- paste("evenness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
even_Stat_Seas_KW_try$Fraction <- factor(even_Stat_Seas_KW_try$Fraction,levels=c("B","E"))
even_Stat_Seas_KW_try$Season <- factor(even_Stat_Seas_KW_try$Season,levels=c("Sp","Su","Fa"))
even_Stat_Seas_KW_try$Station <- factor(even_Stat_Seas_KW_try$Station,levels=c("MLB","MM15","MM110"))
even_Stat_Seas_KW_try$Depth_C <- factor(even_Stat_Seas_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_evenness_Seas_Stat_boxplot.pdf", width= 5, height=5, pointsize= 8)
even_Seas_Stat_boxplot <- ggplot(even_Stat_Seas_KW_try, aes(x = Stat_Seas, y = mean))  +
  geom_boxplot(outlier.size = 0)  + ggtitle("") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = even_Stat_Seas_KW_sigs, aes(label = siglabel, x = Stat_Seas, y = 0), size = 2) +
  scale_color_manual(name = "season", values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),
                     labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                     labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
    scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("MLB Sp", "MM15 Sp", "MM110 Sp","MLB Su", "MM15 Su", "MM110 Su","MLB Fa", "MM15 Fa", "MM110 Fa"),
                   labels=c("MLB Sp", "MM15 Sp", "MM110 Sp","MLB Su", "MM15 Su", "MM110 Su","MLB Fa", "MM15 Fa", "MM110 Fa")) + 
  xlab("") + ylab("Simpson' Evenness") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); even_Seas_Stat_boxplot #top, right, bottom, left
dev.off()

## 2. evenness differences based on Fraction and Season
even_Frac_Seas_KW <- kruskal.test(simps_even$mean ~ as.factor(simps_even$Frac_Seas)) 
print(even_Frac_Seas_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
even_Frac_Seas_KW_MC <- kruskalmc(simps_even$mean ~ as.factor(simps_even$Frac_Seas))  ## Defaults to P < 0.05
print(even_Frac_Seas_KW_MC)

# Test to find the letters to represent significance in a plot
even_Frac_Seas_KW_test <- even_Frac_Seas_KW_MC$dif.com$difference # select logical vector
names(even_Frac_Seas_KW_test) <- row.names(even_Frac_Seas_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
even_Frac_Seas_KW_letters <- multcompLetters(even_Frac_Seas_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
even_Frac_Seas_KW_sigs_dataframe <-  data.frame(as.vector(names(even_Frac_Seas_KW_letters$Letters)), as.vector(even_Frac_Seas_KW_letters$Letters))
colnames(even_Frac_Seas_KW_sigs_dataframe) <- c("Frac_Seas", "siglabel")
even_Frac_Seas_KW_try <- merge(simps_even, even_Frac_Seas_KW_sigs_dataframe, by = "Frac_Seas")

even_Frac_Seas_KW_try$Frac_Seas <-factor(even_Frac_Seas_KW_try$Frac_Seas,levels=c( "B Sp", "B Su", "B Fa", "E Sp", "E Su", "E Fa"))

even_Frac_Seas_KW_sigs <- even_Frac_Seas_KW_try %>%
  select(Frac_Seas, Frac_Seas, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("evenness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
even_Frac_Seas_KW_try$Fraction <- factor(even_Frac_Seas_KW_try$Fraction,levels=c("B","E"))
even_Frac_Seas_KW_try$Season <- factor(even_Frac_Seas_KW_try$Season,levels=c("Sp","Su","Fa"))
even_Frac_Seas_KW_try$Station <- factor(even_Frac_Seas_KW_try$Station,levels=c("MLB","MM15","MM110"))
even_Frac_Seas_KW_try$Depth_C <- factor(even_Frac_Seas_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_evenness_Frac_Seas_boxplot.pdf", width= 5, height=5, pointsize= 8)
even_Frac_Seas_boxplot <- ggplot(even_Frac_Seas_KW_try, aes(x = Frac_Seas, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 0)  + ggtitle("") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = even_Frac_Seas_KW_sigs, aes(label = siglabel, x = Frac_Seas, y = 0), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
 scale_x_discrete(breaks=c("B Sp", "B Su", "B Fa", "E Sp", "E Su", "E Fa"),
                   labels=c("FL Sp", "FL Su", "FL Fa", "PA Sp", "PA Su", "PA Fa")) + 
   xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); even_Frac_Seas_boxplot #top, right, bottom, left
dev.off()


## 3. evenness differences based on Station and Season
#  Resort the factors of Season & Fraction object to have primary sorting by station
even_Stat_Seas_KW_try$Stat_Seas <-factor(even_Stat_Seas_KW_try$Stat_Seas,levels=c( "MLB Sp","MLB Su","MLB Fa","MM15 Sp","MM15 Su","MM15 Fa","MM110 Sp","MM110 Su","MM110 Fa"))

even_Stat_Seas_KW_sigs <- even_Stat_Seas_KW_try %>%
  select(Stat_Seas, Stat_Seas, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("evenness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
even_Station_KW_try$Fraction <- factor(even_Station_KW_try$Fraction,levels=c("B","E"))
even_Station_KW_try$Season <- factor(even_Station_KW_try$Season,levels=c("Sp","Su","Fa"))
even_Station_KW_try$Station <- factor(even_Station_KW_try$Station,levels=c("MLB","MM15","MM110"))
even_Station_KW_try$Depth_C <- factor(even_Station_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_evenness_Stat_Seas_boxplot.pdf", width= 5, height=5, pointsize= 8)
even_Stat_Seas_boxplot <- ggplot(even_Stat_Seas_KW_try, aes(x = Stat_Seas, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 0)  + ggtitle("") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = even_Stat_Seas_KW_sigs, aes(label = siglabel, x = Stat_Seas, y = 0), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("MLB Sp", "MM15 Sp", "MM110 Sp","MLB Su", "MM15 Su", "MM110 Su","MLB Fa", "MM15 Fa", "MM110 Fa"),
                   labels=c("MLB Sp", "MM15 Sp", "MM110 Sp","MLB Su", "MM15 Su", "MM110 Su","MLB Fa", "MM15 Fa", "MM110 Fa")) +   xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); even_Stat_Seas_boxplot #top, right, bottom, left
dev.off()


## 4. evenness differences based on Depth_C and Season
#simps_even_off <- simps_even[simps_even$Station == "MM110" & simps_even$Season == "Su", ]

even_Depth_Seas_KW <- kruskal.test(simps_even$mean ~ as.factor(simps_even$Depth_Seas)) 
print(even_Depth_Seas_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
even_Depth_Seas_KW_MC <- kruskalmc(simps_even$mean ~ as.factor(simps_even$Depth_Seas))  ## Defaults to P < 0.05
print(even_Depth_Seas_KW_MC)

# Test to find the letters to represent significance in a plot
even_Depth_Seas_KW_test <- even_Depth_Seas_KW_MC$dif.com$difference # select logical vector
names(even_Depth_Seas_KW_test) <- row.names(even_Depth_Seas_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
even_Depth_Seas_KW_letters <- multcompLetters(even_Depth_Seas_KW_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
even_Depth_Seas_KW_sigs_dataframe <-  data.frame(as.vector(names(even_Depth_Seas_KW_letters$Letters)), as.vector(even_Depth_Seas_KW_letters$Letters))
colnames(even_Depth_Seas_KW_sigs_dataframe) <- c("Depth_Seas", "siglabel")
even_Depth_Seas_KW_try <- merge(simps_even, even_Depth_Seas_KW_sigs_dataframe, by = "Depth_Seas")

even_Depth_Seas_KW_try$Depth_Seas <-factor(even_Depth_Seas_KW_try$Depth_Seas,levels=c( "S Sp", "S Su", "S Fa", "M Su", "D Sp", "D Su", "D Fa"))

even_Depth_Seas_KW_sigs <- even_Depth_Seas_KW_try %>%
  select(Depth_Seas, Depth_Seas, siglabel) %>%
  distinct()

# finally, the plot
alpha_title <- paste("evenness when rarefied at",min(sample_sums(physeq_dupmerge)),"reads")
even_Depth_Seas_KW_try$Fraction <- factor(even_Depth_Seas_KW_try$Fraction,levels=c("B","E"))
even_Depth_Seas_KW_try$Season <- factor(even_Depth_Seas_KW_try$Season,levels=c("Sp","Su","Fa"))
even_Depth_Seas_KW_try$Station <- factor(even_Depth_Seas_KW_try$Station,levels=c("MLB","MM15","MM110"))
even_Depth_Seas_KW_try$Depth_C <- factor(even_Depth_Seas_KW_try$Depth_C,levels=c("S","M","D"))

pdf(file="Figures/Rplot_evenness_Depth_Seas_boxplot.pdf", width= 5, height=5, pointsize= 8)
even_Depth_Seas_boxplot <- ggplot(even_Depth_Seas_KW_try, aes(x = Depth_Seas, y = mean))  +
  geom_boxplot(aes(color=Season),outlier.size = 0)  + ggtitle("") +
  geom_point(aes(color=Season,shape = Station,size = Fraction,fill=Depth_Seas), position = position_dodge(0.8)) +
  geom_text(data = even_Depth_Seas_KW_sigs, aes(label = siglabel, x = Depth_Seas, y = 0), size = 2) +
  scale_color_manual(name = "season", breaks = c("Sp","Su","Fa"),values = c("Sp" = "darkgreen", "Su" = "red", "Fa" = "mediumblue"),labels = c("Spring",  "Summer", "Fall")) +
  scale_shape_manual(name = "station", values = c("MLB" = 21, "MM15" = 22, "MM110" = 24),
                      labels = c("Estuary", "Near",  "Off")) + 
  scale_size_manual(name = "fraction", breaks = c("B","E"), 
                     labels = c("Free","Particle"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  scale_x_discrete(breaks=c("S Sp", "S Su", "S Fa", "M Su", "D Sp", "D Su", "D Fa"),
                   labels=c("S Sp", "S Su", "S Fa", "M Su", "D Sp", "D Su", "D Fa")) + 
  xlab("") + ylab("") + theme_bw() + 
  theme(plot.title = element_text(face="bold", size=7),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=7),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=7),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 0.5, size=7),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=7),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); even_Depth_Seas_boxplot #top, right, bottom, left
dev.off()
```

```{r final figures}
library(grid)
library(gridExtra)

pdf(file="Figures/Rplot_AlphaDiv_boxplot.pdf", width= 7, height=5, pointsize= 8, colormodel="cmyk", useDingbats = FALSE)
grid.arrange(rich_Season_boxplot,rich_Station_boxplot,rich_Fraction_boxplot,rich_Depth_C_boxplot,even_Season_boxplot,even_Station_boxplot,even_Fraction_boxplot,even_Depth_C_boxplot,ncol=4)
dev.off()

pdf(file="Figures/Rplot_AlphaDiv_multifact_boxplot.pdf", width= 7, height=5, pointsize= 8, colormodel="cmyk", useDingbats = FALSE)
grid.arrange(rich_Seas_Stat_boxplot,rich_Stat_Seas_boxplot,rich_Frac_Seas_boxplot,rich_Depth_Seas_boxplot,even_Seas_Stat_boxplot,even_Stat_Seas_boxplot,even_Frac_Seas_boxplot,even_Depth_Seas_boxplot,ncol=4)
dev.off()
```

### Figure 3 + Table X: PCoA, PERMANOVA
```{r adonis}
# Calculate bray curtis
physeq.bray <- phyloseq::distance(physeq = physeq.scale, method = "bray")

### Let's run ADONIS using Michelle Berry's MicrobeMiseq function --- FOR TABLE 2
phyloseq_to_adonis(physeq.scale, distmat = physeq.bray, dist = "bray","Fraction")
phyloseq_to_adonis(physeq.scale, distmat = physeq.bray, dist = "bray","Season")
phyloseq_to_adonis(physeq.scale, distmat = physeq.bray, dist = "bray","Station")
phyloseq_to_adonis(physeq.scale, distmat = physeq.bray, dist = "bray","Depth_C")
phyloseq_to_adonis(physeq.scale, distmat = physeq.bray, dist = "bray","DayNight")
phyloseq_to_adonis(physeq.scale, distmat = physeq.bray, dist = "bray","Lake")
formula<-c("Fraction","Season","Lake","Station","Depth_C","DayNight")
phyloseq_to_adonis(physeq.scale, distmat = physeq.bray, dist = "bray",formula)

# Prune out rare taxa
# Keep OTUs above average relative abundance of 7 reads (0.01%)
a <- apply(otu_table(physeq.scale), MARGIN = 1, function(x) {mean(x)} )
keep <- a > 7
physeq.prune <- prune_taxa(keep, physeq.scale)
ntaxa(physeq.prune) 

# Calculate Sorensen
physeq.sor <- phyloseq::distance(physeq = physeq.prune, method = "bray", binary=TRUE)

### Let's run ADONIS using Michelle Berry's MicrobeMiseq function --- FOR TABLE 2
phyloseq_to_adonis(physeq.scale, distmat = physeq.sor, dist = "sorensen","Fraction")
phyloseq_to_adonis(physeq.scale, distmat = physeq.sor, dist = "sorensen","Season")
phyloseq_to_adonis(physeq.scale, distmat = physeq.sor, dist = "sorensen","Station")
phyloseq_to_adonis(physeq.scale, distmat = physeq.sor, dist = "sorensen","Depth_C")
phyloseq_to_adonis(physeq.scale, distmat = physeq.sor, dist = "sorensen","DayNight")
phyloseq_to_adonis(physeq.scale, distmat = physeq.sor, dist = "sorensen","Lake")
formula<-c("Fraction","Season","Lake","Station","Depth_C","DayNight")
phyloseq_to_adonis(physeq.scale, distmat = physeq.sor, dist = "sorensen",formula)
```

```{r pcoa bray-curtis}
# Betadiv pcoa Bray-Curtis DNA only 
physeq.pcoa <-
  ordinate(
    physeq = physeq.scale,
    method = "PCoA",
    distance = "bray"
  )

physeq.pcoa.vectors <- data.frame(physeq.pcoa$vectors[, 1:4])
physeq.pcoa.vectors$Duplicates <- row.names(physeq.pcoa.vectors)
physeq.pcoa.df <- merge(physeq.pcoa.vectors,data.frame(duplicate_metadata),by="Duplicates")
for(i in 1:length(physeq.pcoa.df$Season)){
  physeq.pcoa.df$Depth_Seas[i]<-paste(as.character(physeq.pcoa.df$Depth_C[i]),
                                  as.character(physeq.pcoa.df$Season[i]))}
bray_values_DNA <- physeq.pcoa$values
bray_rel_eigens_DNA <- bray_values_DNA$Relative_eig
bray_rel_eigen1_DNA <- bray_rel_eigens_DNA[1]
bray_rel_eigen1_percent_DNA <- round(bray_rel_eigen1_DNA * 100, digits = 1)
bray_rel_eigen2_DNA <- bray_rel_eigens_DNA[2]
bray_rel_eigen2_percent_DNA <- round(bray_rel_eigen2_DNA * 100, digits = 1)
bray_rel_eigen3_DNA <- bray_rel_eigens_DNA[3]
bray_rel_eigen3_percent_DNA <- round(bray_rel_eigen3_DNA * 100, digits = 1)
bray_rel_eigen4_DNA <- bray_rel_eigens_DNA[4]
bray_rel_eigen4_percent_DNA <- round(bray_rel_eigen4_DNA * 100, digits = 1)
bray_axis1_DNA <- paste("PCoA1:",bray_rel_eigen1_percent_DNA,"%")
bray_axis2_DNA <- paste("PCoA2:",bray_rel_eigen2_percent_DNA,"%")
bray_axis3_DNA <- paste("PCoA3:",bray_rel_eigen3_percent_DNA,"%")
bray_axis4_DNA <- paste("PCoA4:",bray_rel_eigen4_percent_DNA,"%")

PCoA_DNA_title <- paste("Bray-Curtis,",ntaxa(physeq.scale),"OTUs")

pdf(file="Figures/Rplot_PCOA12_BC_DNA.pdf", width= 20, height=15, pointsize= 8)
pcoa_BC12_DNA <- ggplot(physeq.pcoa.df, aes(Axis.1, Axis.2, color = Season, shape = Station, size = Fraction, fill = Depth_Seas)) +
  xlab(bray_axis1_DNA) + ylab(bray_axis2_DNA) + ggtitle(PCoA_DNA_title) +
  geom_point(alpha=0.9) + theme_bw() + #ylim(1, -1) + xlim(1, -1) +
  scale_color_manual(name = "season", breaks=c("Sp", "Su",  "Fa"),
                     labels = c("Spring", "Summer",  "Fall"), 
                     values = c("Fa" = "mediumblue", "Sp" = "darkgreen", "Su" = "red")) +
  scale_shape_manual(name = "station", breaks = c("MLB", "MM15",  "MM110"), 
                     labels = c("MLB", "M15",  "M110"),
                     values = c("MLB" = 21, "MM15" = 22, "MM110" = 24)) +
  scale_size_manual(name = "fraction", breaks = c("E", "B"), 
                     labels = c("Particle", "Free"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  theme(plot.title = element_text(face="bold", size=8),  #Set the plot title
        axis.text.x = element_text(colour="black", vjust=0.5, size=8), 
        axis.text.y = element_text(colour="black", vjust=0.5, size=8),
        axis.title.x = element_text(face="bold", size=8),
        axis.title.y = element_text(face="bold", size=8),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        ###LEGEND TOP RIGHT CORNER
        legend.position = "none"); pcoa_BC12_DNA
  dev.off()

pdf(file="Figures/Rplot_PCOA13_BC_DNA.pdf", width= 20, height=15, pointsize= 8)
pcoa_BC13_DNA <- ggplot(physeq.pcoa.df, aes(Axis.1, Axis.3, color = Season, shape = Station, size = Fraction, fill = Depth_Seas)) +
  xlab(bray_axis1_DNA) + ylab(bray_axis3_DNA) + ggtitle(PCoA_DNA_title) +
  geom_point(alpha=0.9) + theme_bw() + #ylim(1, -1) + xlim(1, -1) +
  scale_color_manual(name = "season", breaks=c("Sp", "Su",  "Fa"),
                     labels = c("Spring", "Summer",  "Fall"), 
                     values = c("Fa" = "mediumblue", "Sp" = "darkgreen", "Su" = "red")) +
  scale_shape_manual(name = "station", breaks = c("MLB", "MM15",  "MM110"), 
                     labels = c("MLB", "M15",  "M110"),
                     values = c("MLB" = 21, "MM15" = 22, "MM110" = 24)) +
  scale_size_manual(name = "fraction", breaks = c("E", "B"), 
                     labels = c("Particle", "Free"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  theme(plot.title = element_text(face="bold", size=8),  #Set the plot title
        axis.text.x = element_text(colour="black", vjust=0.5, size=8), 
        axis.text.y = element_text(colour="black", vjust=0.5, size=8),
        axis.title.x = element_text(face="bold", size=8),
        axis.title.y = element_text(face="bold", size=8),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        ###LEGEND TOP RIGHT CORNER
        legend.position = "none"); pcoa_BC13_DNA
  dev.off()

```

```{r pcoa bray curtis after removal otu <0.01% on average}
# Prune out rare taxa
# Keep OTUs above average relative abundance of 7 reads (0.01%)
a <- apply(otu_table(physeq.scale), MARGIN = 1, function(x) {mean(x)} )
keep <- a > 7
physeq.prune <- prune_taxa(keep, physeq.scale)
ntaxa(physeq.prune) 

# Betadiv pcoa Bray-Curtis DNA only 
physeq.pcoa.prune <-
  ordinate(
    physeq = physeq.prune,
    method = "PCoA",
    distance = "bray"
  )

physeq.pcoa.prune.vectors <- data.frame(physeq.pcoa.prune$vectors[, 1:4])
physeq.pcoa.prune.vectors$Duplicates <- row.names(physeq.pcoa.prune.vectors)
physeq.pcoa.prune.df <- merge(physeq.pcoa.prune.vectors,data.frame(duplicate_metadata),by="Duplicates")
for(i in 1:length(physeq.pcoa.prune.df$Season)){
  physeq.pcoa.prune.df$Depth_Seas[i]<-paste(as.character(physeq.pcoa.prune.df$Depth_C[i]),
                                  as.character(physeq.pcoa.prune.df$Season[i]))}
bray_values_DNA <- physeq.pcoa.prune$values
bray_rel_eigens_DNA <- bray_values_DNA$Relative_eig
bray_rel_eigen1_DNA <- bray_rel_eigens_DNA[1]
bray_rel_eigen1_percent_DNA <- round(bray_rel_eigen1_DNA * 100, digits = 1)
bray_rel_eigen2_DNA <- bray_rel_eigens_DNA[2]
bray_rel_eigen2_percent_DNA <- round(bray_rel_eigen2_DNA * 100, digits = 1)
bray_rel_eigen3_DNA <- bray_rel_eigens_DNA[3]
bray_rel_eigen3_percent_DNA <- round(bray_rel_eigen3_DNA * 100, digits = 1)
bray_rel_eigen4_DNA <- bray_rel_eigens_DNA[4]
bray_rel_eigen4_percent_DNA <- round(bray_rel_eigen4_DNA * 100, digits = 1)
bray_axis1_DNA <- paste("PCoA1:",bray_rel_eigen1_percent_DNA,"%")
bray_axis2_DNA <- paste("PCoA2:",bray_rel_eigen2_percent_DNA,"%")
bray_axis3_DNA <- paste("PCoA3:",bray_rel_eigen3_percent_DNA,"%")
bray_axis4_DNA <- paste("PCoA4:",bray_rel_eigen4_percent_DNA,"%")

PCoA_DNA_title <- paste("Bray-Curtis,",ntaxa(physeq.prune),"OTUs")

pdf(file="Figures/Rplot_PCOA12_BC.PRUNE_DNA.pdf", width= 20, height=15, pointsize= 8)
pcoa_BC12.PRUNE_DNA <- ggplot(physeq.pcoa.prune.df, aes(Axis.1, Axis.2, color = Season, shape = Station, size = Fraction, fill = Depth_Seas)) +
  xlab(bray_axis1_DNA) + ylab(bray_axis2_DNA) + ggtitle(PCoA_DNA_title) +
  geom_point(alpha=0.9) + theme_bw() + #ylim(1, -1) + xlim(1, -1) +
  scale_color_manual(name = "season", breaks=c("Sp", "Su",  "Fa"),
                     labels = c("Spring", "Summer",  "Fall"), 
                     values = c("Fa" = "mediumblue", "Sp" = "darkgreen", "Su" = "red")) +
  scale_shape_manual(name = "station", breaks = c("MLB", "MM15",  "MM110"), 
                     labels = c("MLB", "M15",  "M110"),
                     values = c("MLB" = 21, "MM15" = 22, "MM110" = 24)) +
  scale_size_manual(name = "fraction", breaks = c("E", "B"), 
                     labels = c("Particle", "Free"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  theme(plot.title = element_text(face="bold", size=8),  #Set the plot title
        axis.text.x = element_text(colour="black", vjust=0.5, size=8), 
        axis.text.y = element_text(colour="black", vjust=0.5, size=8),
        axis.title.x = element_text(face="bold", size=8),
        axis.title.y = element_text(face="bold", size=8),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        ###LEGEND TOP RIGHT CORNER
        legend.position = "none"); pcoa_BC12.PRUNE_DNA
  dev.off()

pdf(file="Figures/Rplot_PCOA13_BC_DNA.pdf", width= 20, height=15, pointsize= 8)
pcoa_BC13.PRUNE_DNA <- ggplot(physeq.pcoa.prune.df, aes(Axis.1, Axis.3, color = Season, shape = Station, size = Fraction, fill = Depth_Seas)) +
  xlab(bray_axis1_DNA) + ylab(bray_axis3_DNA) + ggtitle(PCoA_DNA_title) +
  geom_point(alpha=0.9) + theme_bw() + #ylim(1, -1) + xlim(1, -1) +
  scale_color_manual(name = "season", breaks=c("Sp", "Su",  "Fa"),
                     labels = c("Spring", "Summer",  "Fall"), 
                     values = c("Fa" = "mediumblue", "Sp" = "darkgreen", "Su" = "red")) +
  scale_shape_manual(name = "station", breaks = c("MLB", "MM15",  "MM110"), 
                     labels = c("MLB", "M15",  "M110"),
                     values = c("MLB" = 21, "MM15" = 22, "MM110" = 24)) +
  scale_size_manual(name = "fraction", breaks = c("E", "B"), 
                     labels = c("Particle", "Free"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  theme(plot.title = element_text(face="bold", size=8),  #Set the plot title
        axis.text.x = element_text(colour="black", vjust=0.5, size=8), 
        axis.text.y = element_text(colour="black", vjust=0.5, size=8),
        axis.title.x = element_text(face="bold", size=8),
        axis.title.y = element_text(face="bold", size=8),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        ###LEGEND TOP RIGHT CORNER
        legend.position = "none"); pcoa_BC13.PRUNE_DNA
  dev.off()

```

```{r pcoa Sørensen}
# Betadiv pcoa Sørensen DNA only 
physeq.pcoa.sor <-
  ordinate(
    physeq = physeq.scale,
    method = "PCoA",
    distance = "bray",
    binary = TRUE
  )

physeq.pcoa.sor.vectors <- data.frame(physeq.pcoa.sor$vectors[, 1:4])
physeq.pcoa.sor.vectors$Duplicates <- row.names(physeq.pcoa.sor.vectors)
physeq.pcoa.sor.df <- merge(physeq.pcoa.sor.vectors,data.frame(duplicate_metadata),by="Duplicates")
for(i in 1:length(physeq.pcoa.sor.df$Season)){
  physeq.pcoa.sor.df$Depth_Seas[i]<-paste(as.character(physeq.pcoa.sor.df$Depth_C[i]),
                                  as.character(physeq.pcoa.sor.df$Season[i]))}
sor_values_DNA <- physeq.pcoa.sor$values
sor_rel_eigens_DNA <- sor_values_DNA$Relative_eig
sor_rel_eigen1_DNA <- sor_rel_eigens_DNA[1]
sor_rel_eigen1_percent_DNA <- round(sor_rel_eigen1_DNA * 100, digits = 1)
sor_rel_eigen2_DNA <- sor_rel_eigens_DNA[2]
sor_rel_eigen2_percent_DNA <- round(sor_rel_eigen2_DNA * 100, digits = 1)
sor_rel_eigen3_DNA <- sor_rel_eigens_DNA[3]
sor_rel_eigen3_percent_DNA <- round(sor_rel_eigen3_DNA * 100, digits = 1)
sor_rel_eigen4_DNA <- sor_rel_eigens_DNA[4]
sor_rel_eigen4_percent_DNA <- round(sor_rel_eigen4_DNA * 100, digits = 1)
sor_axis1_DNA <- paste("PCoA1:",sor_rel_eigen1_percent_DNA,"%")
sor_axis2_DNA <- paste("PCoA2:",sor_rel_eigen2_percent_DNA,"%")
sor_axis3_DNA <- paste("PCoA3:",sor_rel_eigen3_percent_DNA,"%")
sor_axis4_DNA <- paste("PCoA4:",sor_rel_eigen4_percent_DNA,"%")

PCoA_Sor_DNA_title <- paste("Sørensen,",ntaxa(physeq.scale),"OTUs")

pdf(file="Figures/Rplot_PCOA12_SOR_DNA.pdf", width= 20, height=15, pointsize= 8)
pcoa_SOR12_DNA <- ggplot(physeq.pcoa.sor.df, aes(-Axis.1, -Axis.2, color = Season, shape = Station, size = Fraction, fill = Depth_Seas)) +
  xlab(sor_axis1_DNA) + ylab(sor_axis2_DNA) + ggtitle(PCoA_Sor_DNA_title) +
  geom_point(alpha=0.9) + theme_bw() + #ylim(1, -1) + xlim(1, -1) +
  scale_color_manual(name = "season", breaks=c("Sp", "Su",  "Fa"),
                     labels = c("Spring", "Summer",  "Fall"), 
                     values = c("Fa" = "mediumblue", "Sp" = "darkgreen", "Su" = "red")) +
  scale_shape_manual(name = "station", breaks = c("MLB", "MM15",  "MM110"), 
                     labels = c("MLB", "M15",  "M110"),
                     values = c("MLB" = 21, "MM15" = 22, "MM110" = 24)) +
  scale_size_manual(name = "fraction", breaks = c("E", "B"), 
                     labels = c("Particle", "Free"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  theme(plot.title = element_text(face="bold", size=8),  #Set the plot title
        axis.text.x = element_text(colour="black", vjust=0.5, size=8), 
        axis.text.y = element_text(colour="black", vjust=0.5, size=8),
        axis.title.x = element_text(face="bold", size=8),
        axis.title.y = element_text(face="bold", size=8),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        ###LEGEND TOP RIGHT CORNER
        legend.position = "none"); pcoa_SOR12_DNA
  dev.off()

pdf(file="Figures/Rplot_PCOA13_SOR_DNA.pdf", width= 20, height=15, pointsize= 8)
pcoa_SOR13_DNA <- ggplot(physeq.pcoa.sor.df, aes(-Axis.1, -Axis.3, color = Season, shape = Station, size = Fraction, fill = Depth_Seas)) +
  xlab(sor_axis1_DNA) + ylab(sor_axis3_DNA) + ggtitle(PCoA_Sor_DNA_title) +
  geom_point(alpha=0.9) + theme_bw() + #ylim(1, -1) + xlim(1, -1) +
  scale_color_manual(name = "season", breaks=c("Sp", "Su",  "Fa"),
                     labels = c("Spring", "Summer",  "Fall"), 
                     values = c("Fa" = "mediumblue", "Sp" = "darkgreen", "Su" = "red")) +
  scale_shape_manual(name = "station", breaks = c("MLB", "MM15",  "MM110"), 
                     labels = c("MLB", "M15",  "M110"),
                     values = c("MLB" = 21, "MM15" = 22, "MM110" = 24)) +
  scale_size_manual(name = "fraction", breaks = c("E", "B"), 
                     labels = c("Particle", "Free"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  theme(plot.title = element_text(face="bold", size=8),  #Set the plot title
        axis.text.x = element_text(colour="black", vjust=0.5, size=8), 
        axis.text.y = element_text(colour="black", vjust=0.5, size=8),
        axis.title.x = element_text(face="bold", size=8),
        axis.title.y = element_text(face="bold", size=8),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        ###LEGEND TOP RIGHT CORNER
        legend.position = "none"); pcoa_SOR13_DNA
  dev.off()

```

```{r pcoa Sørensen after removal of OTU < 0.01% on average}
# Prune out rare taxa
# Keep OTUs above average relative abundance of 7 reads (0.01%)
a <- apply(otu_table(physeq.scale), MARGIN = 1, function(x) {mean(x)} )
keep <- a > 7
physeq.prune <- prune_taxa(keep, physeq.scale)
ntaxa(physeq.prune) 

# Betadiv pcoa Sørensen DNA only 
physeq.pcoa.prune.sor <-
  ordinate(
    physeq = physeq.prune,
    method = "PCoA",
    distance = "bray",
    binary = TRUE
  )

physeq.pcoa.prune.sor.vectors <- data.frame(physeq.pcoa.prune.sor$vectors[, 1:4])
physeq.pcoa.prune.sor.vectors$Duplicates <- row.names(physeq.pcoa.prune.sor.vectors)
physeq.pcoa.prune.sor.df <- merge(physeq.pcoa.prune.sor.vectors,data.frame(duplicate_metadata),by="Duplicates")
for(i in 1:length(physeq.pcoa.prune.sor.df$Season)){
  physeq.pcoa.prune.sor.df$Depth_Seas[i]<-paste(as.character(physeq.pcoa.prune.sor.df$Depth_C[i]),
                                  as.character(physeq.pcoa.prune.sor.df$Season[i]))}
sor_values_DNA <- physeq.pcoa.prune.sor$values
sor_rel_eigens_DNA <- sor_values_DNA$Relative_eig
sor_rel_eigen1_DNA <- sor_rel_eigens_DNA[1]
sor_rel_eigen1_percent_DNA <- round(sor_rel_eigen1_DNA * 100, digits = 1)
sor_rel_eigen2_DNA <- sor_rel_eigens_DNA[2]
sor_rel_eigen2_percent_DNA <- round(sor_rel_eigen2_DNA * 100, digits = 1)
sor_rel_eigen3_DNA <- sor_rel_eigens_DNA[3]
sor_rel_eigen3_percent_DNA <- round(sor_rel_eigen3_DNA * 100, digits = 1)
sor_rel_eigen4_DNA <- sor_rel_eigens_DNA[4]
sor_rel_eigen4_percent_DNA <- round(sor_rel_eigen4_DNA * 100, digits = 1)
sor_axis1_DNA <- paste("PCoA1:",sor_rel_eigen1_percent_DNA,"%")
sor_axis2_DNA <- paste("PCoA2:",sor_rel_eigen2_percent_DNA,"%")
sor_axis3_DNA <- paste("PCoA3:",sor_rel_eigen3_percent_DNA,"%")
sor_axis4_DNA <- paste("PCoA4:",sor_rel_eigen4_percent_DNA,"%")

PCoA_Sor_DNA_title <- paste("Sørensen,",ntaxa(physeq.prune),"OTUs")

pdf(file="Figures/Rplot_PCOA12_SOR.PRUNE_DNA.pdf", width= 20, height=15, pointsize= 8)
pcoa_SOR12.PRUNE_DNA <- ggplot(physeq.pcoa.prune.sor.df, aes(-Axis.1, -Axis.2, color = Season, shape = Station, size = Fraction, fill = Depth_Seas)) +
  xlab(sor_axis1_DNA) + ylab(sor_axis2_DNA) + ggtitle(PCoA_Sor_DNA_title) +
  geom_point(alpha=0.9) + theme_bw() + #ylim(1, -1) + xlim(1, -1) +
  scale_color_manual(name = "season", breaks=c("Sp", "Su",  "Fa"),
                     labels = c("Spring", "Summer",  "Fall"), 
                     values = c("Fa" = "mediumblue", "Sp" = "darkgreen", "Su" = "red")) +
  scale_shape_manual(name = "station", breaks = c("MLB", "MM15",  "MM110"), 
                     labels = c("MLB", "M15",  "M110"),
                     values = c("MLB" = 21, "MM15" = 22, "MM110" = 24)) +
  scale_size_manual(name = "fraction", breaks = c("E", "B"), 
                     labels = c("Particle", "Free"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  theme(plot.title = element_text(face="bold", size=8),  #Set the plot title
        axis.text.x = element_text(colour="black", vjust=0.5, size=8), 
        axis.text.y = element_text(colour="black", vjust=0.5, size=8),
        axis.title.x = element_text(face="bold", size=8),
        axis.title.y = element_text(face="bold", size=8),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        ###LEGEND TOP RIGHT CORNER
        legend.position = "none"); pcoa_SOR12.PRUNE_DNA
  dev.off()

pdf(file="Figures/Rplot_PCOA13_SOR.PRUNE_DNA.pdf", width= 20, height=15, pointsize= 8)
pcoa_SOR13.PRUNE_DNA <- ggplot(physeq.pcoa.prune.sor.df, aes(-Axis.1, -Axis.3, color = Season, shape = Station, size = Fraction, fill = Depth_Seas)) +
  xlab(sor_axis1_DNA) + ylab(sor_axis3_DNA) + ggtitle(PCoA_Sor_DNA_title) +
  geom_point(alpha=0.9) + theme_bw() + #ylim(1, -1) + xlim(1, -1) +
  scale_color_manual(name = "season", breaks=c("Sp", "Su",  "Fa"),
                     labels = c("Spring", "Summer",  "Fall"), 
                     values = c("Fa" = "mediumblue", "Sp" = "darkgreen", "Su" = "red")) +
  scale_shape_manual(name = "station", breaks = c("MLB", "MM15",  "MM110"), 
                     labels = c("MLB", "M15",  "M110"),
                     values = c("MLB" = 21, "MM15" = 22, "MM110" = 24)) +
  scale_size_manual(name = "fraction", breaks = c("E", "B"), 
                     labels = c("Particle", "Free"),
                     values = c("B" = 1, "E" = 2)) +
  scale_fill_manual(name = "depth", breaks = c("D Sp", "D Su", "D Fa", "M Su", "S Sp", "S Su", "S Fa"), 
                     labels = c("Deep","Deep","Deep", "Chla max", "Surface", "Surface", "Surface"),
                     values = c("D Fa" = "mediumblue", "D Sp" = "darkgreen", "D Su" = "red", "M Su" = "orange", "S Sp" = "white", "S Su" = "white", "S Fa" = "white")) +
  theme(plot.title = element_text(face="bold", size=8),  #Set the plot title
        axis.text.x = element_text(colour="black", vjust=0.5, size=8), 
        axis.text.y = element_text(colour="black", vjust=0.5, size=8),
        axis.title.x = element_text(face="bold", size=8),
        axis.title.y = element_text(face="bold", size=8),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        ###LEGEND TOP RIGHT CORNER
        legend.position = "none"); pcoa_SOR13.PRUNE_DNA
  dev.off()

```

```{r final figures}
library(grid)
library(gridExtra)

pdf(file="Figures/Rplot_BetaDiv_pcoa.pdf", width= 7, height=3.5, pointsize= 8, colormodel="cmyk", useDingbats = FALSE)
grid.arrange(pcoa_BC12_DNA,pcoa_SOR12_DNA,ncol=2)
dev.off()

pdf(file="Figures/Rplot_BetaDiv_0.01_pcoa.pdf", width= 7, height=3.5, pointsize= 8, colormodel="cmyk", useDingbats = FALSE)
grid.arrange(pcoa_BC12.PRUNE_DNA,pcoa_SOR12.PRUNE_DNA,ncol=2)
dev.off()
```


##Figure 3: cyano, mixo, hetero
```{r barplots}

#Replaces phylum info with functional group information based on Phylum and Family information
physeq.fg <- physeq
phy <- data.frame(tax_table(physeq.fg))
Phylum <- as.character(phy$Phylum)
Family <- as.character(phy$Family)

#Include bacteriochlorophyl-containing organisms as per http://mmbr.asm.org/content/62/3/695.abstract?ijkey=87bc0335b3da193292146dd6302d0808bb623bda&keytype2=tf_ipsecsha as putative AAP mixotrophs (aerobic anoxygenic photoheterotrophs) and http://aem.asm.org/content/72/12/7431.full
#Roseomonas: http://www.sciencedirect.com/science/article/pii/S0723202009001271

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] != "Cyanobacteria"){
    if (Family[i] == "Rhodobacteraceae"){
      Phylum[i] <- "Putative_mixotroph"       
    } else {
      Phylum[i] <- "Putative_heterotroph"
    }
  } 
}

phy$Phylum <- Phylum
t <- tax_table(as.matrix(phy))
tax_table(physeq.fg) <- t

#merge by combining diel samples at same station, then scale
physeq.fg <- merge_samples(physeq.fg,"Duplicates_nodiel")

# Weird bug, set taxa to rows
if (!taxa_are_rows(physeq.fg)) {
  otu_table(physeq.fg) <- t(otu_table(physeq.fg))
}

## Fix metadata of merged duplicates 
# Grab unique lines (first entry) from inland phyloseq object metadata
unique_idx <- !(duplicated(sample_data(physeq)$Duplicates_nodiel))
fg_metadata <- sample_data(physeq)[unique_idx, ]
# Assign new metadata to phyloseq object
sample_data(physeq.fg) <- fg_metadata
physeq.fg <- prune_taxa(taxa_sums(physeq.fg) > 0, physeq.fg)

#check number of reads in each sample, differences in count are in part due differet numbers of chlorophyl reads depending on sample
sampsums <- data.frame(sums = sample_sums(physeq.fg))
ggplot(sampsums, aes(x = sums)) + geom_histogram(binwidth = 10000)
min(sampsums)
max(sampsums)

# Scales reads to smallest library size which is 100,766
physeq.fg.scale <- scale_reads(physeq.fg, min(sample_sums(physeq.fg)))
physeq.fg.scale <- prune_taxa(taxa_sums(physeq.fg.scale) > 0, physeq.fg.scale)

# Keep OTUs above average relative abundance of 10 reads (0.01% after scaling) 
mean(sample_sums(physeq.fg.scale))

a <- apply(otu_table(physeq.fg.scale), MARGIN = 1, function(x) {mean(x)} )
keep <- a > 10
physeq.fg.scale.prune <- prune_taxa(keep, physeq.fg.scale)
ntaxa(physeq.fg.scale.prune) 

View(tax_table(physeq.fg.scale.prune))


### determine how many sequences included for each sample when only considering abundant OTUs
relab_toptaxa <- sample_sums(physeq.fg.scale.prune)/sample_sums(physeq.fg.scale)*100
mean(relab_toptaxa)
min(relab_toptaxa)
max(relab_toptaxa)
median(relab_toptaxa)
sd(relab_toptaxa)
relab_toptaxa <- data.frame(relab_toptaxa)
ggplot(relab_toptaxa, aes(x = relab_toptaxa)) + geom_histogram(binwidth = .05)

# Now generate data frame in long format (the output from miseqR.R transform_and_melt function) 
# and produce a stacked barplot of community composition.
fg_relab<-transform_and_melt(physeq.fg.scale.prune,"Phylum",0.01)

fg.colors <- c("chartreuse3","darkred", "steelblue4")

fg_relab$Season <- factor(fg_relab$Season, levels = c("Sp","Su","Fa"))
fg_relab$Depth_C <- factor(fg_relab$Depth_C, levels = c("S","M","D"))

# Plot
ggplot(fg_relab,aes(x = Station, y = Abundance, fill=Taxonomy)) + 
  facet_grid(Depth_C+Fraction~Season, scales = "free_y") +
  geom_bar(stat = "identity") +
  geom_bar(stat = "identity",position = "fill",colour = "black",show.legend = FALSE) +
  scale_fill_manual(values = fg.colors) +
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance\n") +
  theme(axis.title.x =element_blank(),strip.text =element_text(size = 14)
)


```

## Figure 4: differentially abundant taxa (PA and FL separately) TO DO: deseqs, then plot most abundant ones with largest effects size as in Michelle's ISMEJ paper?
```{r Deseq analysis MLB vs LM}
mean(sample_sums(physeq_dupmerge))
# Keep OTUs above average relative abundance of 33 reads (0.01% before scaling) for use with deseq 
a <- apply(otu_table(physeq_dupmerge), MARGIN = 1, function(x) {mean(x)} )
keep <- a > 330
physeq_dupmerge.prune <- prune_taxa(keep, physeq_dupmerge)
ntaxa(physeq_dupmerge.prune) 

# Convert to deseq object
physeq_dupmerge.prune.deseq <- phyloseq_to_deseq2(physeq_dupmerge.prune, ~Season+Fraction+Depth_C+Lake)
physeq_dupmerge.prune.deseq = DESeq(physeq_dupmerge.prune.deseq, test = "Wald", fitType = "parametric", betaPrior=FALSE)

# Summarize results
res = data.frame(results(physeq_dupmerge.prune.deseq, cooksCutoff = FALSE))
res = data.frame(OTU = row.names(res), res)
alpha = 0.01
sigtab <-
  res %>%
    filter(padj < alpha) 
sigtab.LMvsMLB <- cbind(sigtab, 
  as(tax_table(physeq_dupmerge.prune)[sigtab$OTU, ], "matrix"))

# Change genus levels so organized by phylum
sigtab.LMvsMLB <- arrange(sigtab.LMvsMLB, Phylum)
sigtab.LMvsMLB$Genus <- factor(sigtab.LMvsMLB$Genus, 
  levels = unique(sigtab.LMvsMLB$Genus))
# Plot
ggplot(sigtab.LMvsMLB, 
  aes(x = Phylum, y = log2FoldChange, color = Phylum)) + 
  geom_point() + 
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 10),
    legend.position = "none"
  ) 


phylum.colors<-group.colors

# deseq without pruning low abundance - otu level - per fraction - PA - no scaling (done by DEseq2!)
raw_merged_PA <- subset_samples(raw_merged_toptaxa, Fraction != "B")
deseq_otu_DNAcDNA_PA_raw <- deSEQ_noprune(raw_merged_PA, ~ DNA)
deseq_otu_DNAcDNA_PA_raw$plotvalue <- as.vector(as.numeric(taxa_sums(raw_merged_PA)/sum(sample_sums(raw_merged_PA))*100)^.25) # abundance scaled to optimize gradient of glyph sizes in image
deseq_otu_DNAcDNA_PA_raw$sig <- as.factor((deseq_otu_DNAcDNA_PA_raw$padj<0.01)*1)
deseq_otu_DNAcDNA_plot_PA_raw <- plot_deSEQ(deseq_otu_DNAcDNA_PA_raw, "OTU cDNA:DNA PA samples")
deseq_otu_DNAcDNA_plot_PA_raw

# deseq without pruning low abundance - otu level - per fraction - FL - no scaling (done by DEseq2!)
raw_merged_FL <- subset_samples(raw_merged_toptaxa, Fraction != "E")
deseq_otu_DNAcDNA_FL_raw <- deSEQ_noprune(raw_merged_FL, ~ DNA)
deseq_otu_DNAcDNA_FL_raw$plotvalue <- as.vector(as.numeric(taxa_sums(raw_merged_FL)/sum(sample_sums(raw_merged_FL))*100)^.25)
deseq_otu_DNAcDNA_FL_raw$sig <- as.factor((deseq_otu_DNAcDNA_FL_raw$padj<0.01)*1)
deseq_otu_DNAcDNA_plot_FL_raw <- plot_deSEQ(deseq_otu_DNAcDNA_FL_raw, "OTU cDNA:DNA FL samples")
deseq_otu_DNAcDNA_plot_FL_raw

#plot together --- FOR FIGURE 4
Phylum_temp <- as.character(deseq_otu_DNAcDNA_PA_raw$Phylum)
for  (i in 1:length(Phylum_temp)){ 
  Phylum_temp[i] = paste(Phylum_temp[i],"PA",sep=" ")
} 
deseq_otu_DNAcDNA_PA_raw$Phylum_plot <- Phylum_temp

Phylum_temp <- as.character(deseq_otu_DNAcDNA_FL_raw$Phylum)
for  (i in 1:length(Phylum_temp)){ 
  Phylum_temp[i] = paste(Phylum_temp[i],"FL",sep=" ")
} 
deseq_otu_DNAcDNA_FL_raw$Phylum_plot <- Phylum_temp

deseq_otu_DNAcDNA_FLandPA_raw<-rbind(deseq_otu_DNAcDNA_FL_raw,deseq_otu_DNAcDNA_PA_raw)

pdf(file="Figures/Rplot_cDvsD_FL_noscale_top200.pdf", width= 10, height=15, pointsize= 8, useDingbats=FALSE)
deseq_otu_DNAcDNA_FLandPA_raw$Phylum_plot <-factor(deseq_otu_DNAcDNA_FLandPA_raw$Phylum_plot,levels=c("Proteobacteria FL","Proteobacteria PA","Alphaproteobacteria FL","Alphaproteobacteria PA","Betaproteobacteria FL","Betaproteobacteria PA","Gammaproteobacteria FL","Gammaproteobacteria PA","Deltaproteobacteria FL","Deltaproteobacteria PA","Bacteroidetes FL","Bacteroidetes PA","Cytophagia FL","Cytophagia PA","Flavobacteria FL","Flavobacteria PA","Sphingobacteria FL","Sphingobacteria PA","Actinobacteria FL","Actinobacteria PA","Cyanobacteria FL","Cyanobacteria PA","Verrucomicrobia FL","Verrucomicrobia PA","Planctomycetes FL","Planctomycetes PA","Chloroflexi FL","Chloroflexi PA","Acidobacteria FL","Acidobacteria PA","Armatimonadetes FL","Armatimonadetes PA","Chlorobi FL","Chlorobi PA","Firmicutes FL","Firmicutes PA","Gemmatimonadetes FL","Gemmatimonadetes PA","unclassified FL","unclassified PA"))
deseq_otu_DNAcDNA_plot_FLandPA_raw <- plot_deSEQ_combo(deseq_otu_DNAcDNA_FLandPA_raw, "OTU cDNA:DNA FL and PA samples")
deseq_otu_DNAcDNA_plot_FLandPA_raw
dev.off()

####pull out most different, highly active OTUs  --- FOR TABLE 3
View(deseq_otu_DNAcDNA_FLandPA_raw)

```

# Overabundant in zm-free lakes
```{r}
# Table
library(knitr)
non.abund <- filter(sigtab.invaded, log2FoldChange > 0)
kable(x = non.abund)
```

# Overabundant in invaded lakes
```{r}
# Table
inv.abund <- filter(sigtab.invaded, log2FoldChange < 0)
kable(x = inv.abund)
```

## top 200 taxa boxplots euclidean distances across seasons (Fig. 6)
```{r top 200 taxa boxplots euclidean distances across seasons}
### determine whether ratios in different habitats becomes increasingly different over seasonal succession

#create dataframe with paired Euclidean dissimilarity and add metadata for compared samples
cDD_norm_euc <- vegdist(t(mf_cD2D_log), method = "euc")  # calculates the Euclidean Distances
cDD_euc <- melt(as.matrix(cDD_norm_euc), varnames = c("samp1", "samp2"))
cDD_euc$season1 <- substr(cDD_euc$samp1,1,2) # Create a new column called season1 with first 2 letters of string
cDD_euc$season2 <- substr(cDD_euc$samp2,1,2) # Create a new column called season2 with first 2 letters of string
cDD_euc$fraction1 <- substr(cDD_euc$samp1,4,4) # Create a new column called fraction1 with letter 4 of string
cDD_euc$fraction2 <- substr(cDD_euc$samp2,4,4) # Create a new column called fraction2 with letter 4 of string
cDD_euc$station1 <- unlist(strsplit(toString(cDD_euc$samp1), "[., ]"))[seq(3,length(unlist(strsplit(toString(cDD_euc$samp1), split="[., ]"))),5)] # Create a new column called station1 with staton name
cDD_euc$station2 <- unlist(strsplit(toString(cDD_euc$samp2), "[., ]"))[seq(3,length(unlist(strsplit(toString(cDD_euc$samp2), split="[., ]"))),5)] # Create a new column called station2 with staton name
cDD_euc$depth1 <- substr(unlist(strsplit(toString(cDD_euc$samp1), "[., ]"))[seq(4,length(unlist(strsplit(toString(cDD_euc$samp1), split="[., ]"))),5)],1,1) # Create a new column called depth1 with depth
cDD_euc$depth2 <- substr(unlist(strsplit(toString(cDD_euc$samp2), "[., ]"))[seq(4,length(unlist(strsplit(toString(cDD_euc$samp2), split="[., ]"))),5)],1,1) # Create a new column called depth2 with depth

#combine category names to group similar comparisons - sorry, clunky code here, but hey, it works!
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp_ML"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp_LM"

cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp_MLB"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp_15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp_110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Sp_FL"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Sp_PA"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Sp_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Sp_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Sp_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Sp_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Sp_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Sp_SurfvsDeep"

cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Su_ML"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Su_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Su_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Su_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Su_LM"

cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Su_MLB"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Su_15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Su_110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Su_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Su_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Su_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Su_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Su_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Su_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Su_FL"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Su_PA"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Su_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Su_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Su_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Su_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Su_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Su"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Su_SurfvsDeep"

cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Fa_ML"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Fa_LM"

cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Fa_MLB"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Fa_15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Fa_110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Fa_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Fa_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Fa_FL"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Fa_PA"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Fa_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Fa_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Fa_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Fa_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Fa_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Fa_SurfvsDeep"

cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLB"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp-Su_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp-Su_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp-Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp-Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp-Su_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp-Su_LM"

cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLB"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp-Su_15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp-Su_110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp-Su_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp-Su_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp-Su_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp-Su_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Sp-Su_FL"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Sp-Su_PA"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Sp-Su_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Sp-Su_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Sp-Su_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Sp-Su_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Sp-Su_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Su"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Sp-Su_SurfvsDeep"

cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLB"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp-Su_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp-Su_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp-Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp-Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp-Su_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp-Su_LM"

cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLB"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp-Su_15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp-Su_110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp-Su_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp-Su_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp-Su_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp-Su_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp-Su_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Sp-Su_FL"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Sp-Su_PA"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Sp-Su_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Sp-Su_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Sp-Su_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Sp-Su_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Sp-Su_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Sp-Su_SurfvsDeep"

cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLB"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp-Fa_LM"

cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLB"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp-Fa_15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp-Fa_110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp-Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp-Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp-Fa_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp-Fa_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Sp-Fa_FL"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Sp-Fa_PA"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Sp-Fa_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Sp-Fa_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Sp-Fa_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Sp-Fa_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Sp-Fa_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Sp"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Sp-Fa_SurfvsDeep"

cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLB"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp-Fa_LM"

cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLB"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Sp-Fa_15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Sp-Fa_110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Sp-Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Sp-Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Sp-Fa_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Sp-Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Sp-Fa_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Sp-Fa_FL"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Sp-Fa_PA"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Sp-Fa_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Sp-Fa_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Sp-Fa_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Sp-Fa_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Sp-Fa_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Sp"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Sp-Fa_SurfvsDeep"

cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLB"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Su-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Su-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Su-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Su-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Su-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Su-Fa_LM"

cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLB"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Su-Fa_15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Su-Fa_110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Su-Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Su-Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Su-Fa_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Su-Fa_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Su-Fa_FL"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Su-Fa_PA"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Su-Fa_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Su-Fa_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Su-Fa_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Su-Fa_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Su-Fa_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Su"&cDD_euc$season2=="Fa"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Su-Fa_SurfvsDeep"

cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLB"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Su-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Su-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Su-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Su-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Su-Fa_LM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLBvsLM"
cDD_euc$complake[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Su-Fa_LM"

cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLB"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="15"]<-"Su-Fa_15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="110"]<-"Su-Fa_110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="15"]<-"Su-Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="MLB"&cDD_euc$station2=="110"]<-"Su-Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="110"]<-"Su-Fa_15vs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="15"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLBvs15"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="MLB"]<-"Su-Fa_MLBvs110"
cDD_euc$compstat[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$station1=="110"&cDD_euc$station2=="15"]<-"Su-Fa_15vs110"

cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="B"]<-"Su-Fa_FL"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="E"]<-"Su-Fa_PA"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="B"&cDD_euc$fraction2=="E"]<-"Su-Fa_FLvsPA"
cDD_euc$compfrac[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$fraction1=="E"&cDD_euc$fraction2=="B"]<-"Su-Fa_FLvsPA"

cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$depth1=="S"&cDD_euc$depth2=="S"]<-"Su-Fa_Surf"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$depth1=="D"&cDD_euc$depth2=="D"]<-"Su-Fa_Deep"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$depth1=="S"&cDD_euc$depth2=="D"]<-"Su-Fa_SurfvsDeep"
cDD_euc$compdepth[cDD_euc$season1=="Fa"&cDD_euc$season2=="Su"&cDD_euc$depth1=="D"&cDD_euc$depth2=="S"]<-"Su-Fa_SurfvsDeep"

#removing self comparisons and calculating average per group, also remove all comparisons between seasons
cDD_euc <- subset(cDD_euc, samp1 != samp2)
cDD_euc <- subset(cDD_euc, season1 == season2)
cDD_euc_frac <- subset(cDD_euc, station1 == station2 & depth1 == depth2 & fraction1 != fraction2)
cDD_euc_stat <- subset(cDD_euc, fraction1 == fraction2 & depth1 == depth2 & station1 != station2)
cDD_euc_lake <- subset(cDD_euc, fraction1 == fraction2 & depth1 == depth2 & complake != "Sp_LM" & complake != "Su_LM" & complake != "Fa_LM")
cDD_euc_depth <- subset(cDD_euc, station1 == station2 & fraction1 == fraction2 & depth1 != depth2)

compfrac_mean <- ddply(cDD_euc_frac, ~compfrac, function(x){data.frame(compfrac_mean = mean(x$value))})
compstat_mean <- ddply(cDD_euc_stat, ~compstat, function(x){data.frame(compstat_mean = mean(x$value))})
complake_mean <- ddply(cDD_euc_lake, ~complake, function(x){data.frame(complake_mean = mean(x$value))})
compdepth_mean <- ddply(cDD_euc_depth, ~compdepth, function(x){data.frame(compdepth_mean = mean(x$value))})

## 1. how does distance between cDNA:DNA in different fractions change over the seasons 
cDD_euc_frac_KW <- kruskal.test(cDD_euc_frac$value ~ as.factor(cDD_euc_frac$compfrac)) 
print(cDD_euc_frac_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
cDD_euc_frac_KW_MC <- kruskalmc(cDD_euc_frac$value ~ as.factor(cDD_euc_frac$compfrac))  ## Defaults to P < 0.05
print(cDD_euc_frac_KW_MC)

# Test to find the letters to represent significance in a plot
euc_frac_test <- cDD_euc_frac_KW_MC$dif.com$difference # select logical vector
names(euc_frac_test) <- row.names(cDD_euc_frac_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
euc_frac_letters <- multcompLetters(euc_frac_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
euc_frac_sigs_dataframe <-  data.frame(as.vector(names(euc_frac_letters$Letters)), as.vector(euc_frac_letters$Letters))
colnames(euc_frac_sigs_dataframe) <- c("compfrac", "siglabel")
euc_frac_try <- merge(cDD_euc_frac, euc_frac_sigs_dataframe, by = "compfrac")

euc_frac_try$compfrac <-factor(euc_frac_try$compfrac,levels=c( "Sp_FLvsPA", "Su_FLvsPA", "Fa_FLvsPA"))

euc_frac_sigs <- euc_frac_try %>%
  select(season1, compfrac, siglabel) %>%
  distinct()

# plot
pdf(file="Figures/Rplot_cDD_euc_frac_boxplot_toptaxa.pdf", width= 5, height=5, pointsize= 8)
euc_frac_boxplot <- ggplot(euc_frac_try, aes(x = compfrac, y = value, fill = compfrac)) + geom_point(size = 2) +
  geom_boxplot() +
  geom_text(data = euc_frac_sigs, aes(label = siglabel, x = compfrac, y = 25), size = 2) +
  scale_fill_manual(name = "", limits=c("Sp_FLvsPA", "Su_FLvsPA", "Fa_FLvsPA"), 
                     values = c("Sp_FLvsPA" = "grey","Su_FLvsPA" = "grey","Fa_FLvsPA" = "grey"))+
  scale_x_discrete(breaks=c("Sp_FLvsPA", "Su_FLvsPA", "Fa_FLvsPA"),
                   labels=c("Spring\n(n=22)", "Summer\n(n=26)", "Fall\n(n=24)")) + 
  xlab("Seasonal distance between cDNA:DNA ratios in FL and PA fractions)") + ylab("Euclidean distance\nlog(cDNA:DNA)") + theme_bw() +
  theme(plot.title = element_text(face="bold", size = 12),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=10),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=10),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 1, size=10),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=10),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); euc_frac_boxplot#top, right, bottom, left
dev.off()

## 2. how does distance between cDNA:DNA in different lakes change over the seasons 
cDD_euc_lake_KW <- kruskal.test(cDD_euc_lake$value ~ as.factor(cDD_euc_lake$complake)) 
print(cDD_euc_lake_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
cDD_euc_lake_KW_MC <- kruskalmc(cDD_euc_lake$value ~ as.factor(cDD_euc_lake$complake))  ## Defaults to P < 0.05
print(cDD_euc_lake_KW_MC)

# Test to find the letters to represent significance in a plot
euc_lake_test <- cDD_euc_lake_KW_MC$dif.com$difference # select logical vector
names(euc_lake_test) <- row.names(cDD_euc_lake_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
euc_lake_letters <- multcompLetters(euc_lake_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
euc_lake_sigs_dataframe <-  data.frame(as.vector(names(euc_lake_letters$Letters)), as.vector(euc_lake_letters$Letters))
colnames(euc_lake_sigs_dataframe) <- c("complake", "siglabel")
euc_lake_try <- merge(cDD_euc_lake, euc_lake_sigs_dataframe, by = "complake")

euc_lake_try$complake <-factor(euc_lake_try$complake,levels=c( "Sp_MLBvsLM", "Su_MLBvsLM", "Fa_MLBvsLM"))

euc_lake_sigs <- euc_lake_try %>%
  select(season1, complake, siglabel) %>%
  distinct()

# plot
pdf(file="Figures/Rplot_cDD_euc_lake_boxplot_toptaxa.pdf", width= 5, height=5, pointsize= 14)
euc_lake_boxplot <- ggplot(euc_lake_try, aes(x = complake, y = value, fill = complake)) + geom_point(size = 2) +
  geom_boxplot() +
  geom_text(data = euc_lake_sigs, aes(label = siglabel, x = complake, y = 25), size = 4) +
  scale_fill_manual(name = "", limits=c("Sp_MLBvsLM", "Su_MLBvsLM", "Fa_MLBvsLM"), 
                     values = c("Sp_MLBvsLM" = "grey","Su_MLBvsLM" = "grey","Fa_MLBvsLM" = "grey"))+
  scale_x_discrete(breaks=c("Sp_MLBvsLM", "Su_MLBvsLM", "Fa_MLBvsLM"),
                   labels=c("Spring\n(n=16)", "Summer\n(n=26)", "Fall\n(n=24)")) + 
  xlab("Seasonal distance between cDNA:DNA ratios in Lake MI and Muskegon Lake)") + ylab("Euclidean distance\nlog(cDNA:DNA)") + theme_bw()  +
  theme(plot.title = element_text(face="bold", size = 12),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=10),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=10),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 1, size=10),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=10),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); euc_lake_boxplot#top, right, bottom, left
dev.off()

## 3. how does distance between cDNA:DNA in different depths change over the seasons 
cDD_euc_depth_KW <- kruskal.test(cDD_euc_depth$value ~ as.factor(cDD_euc_depth$compdepth)) 
print(cDD_euc_depth_KW)  # show Kruskal Wallis result

# Which samples are significantly different from each other?  
cDD_euc_depth_KW_MC <- kruskalmc(cDD_euc_depth$value ~ as.factor(cDD_euc_depth$compdepth))  ## Defaults to P < 0.05
print(cDD_euc_depth_KW_MC)

# Test to find the letters to represent significance in a plot
euc_depth_test <- cDD_euc_depth_KW_MC$dif.com$difference # select logical vector
names(euc_depth_test) <- row.names(cDD_euc_depth_KW_MC$dif.com) # add comparison names

# create a list with "homogenous groups" coded by letter
euc_depth_letters <- multcompLetters(euc_depth_test, compare="<", threshold=0.05, 
                                Letters=c(letters, LETTERS, "."), reversed = FALSE)

#  Extract the values from the multcompLetters object
euc_depth_sigs_dataframe <-  data.frame(as.vector(names(euc_depth_letters$Letters)), as.vector(euc_depth_letters$Letters))
colnames(euc_depth_sigs_dataframe) <- c("compdepth", "siglabel")
euc_depth_try <- merge(cDD_euc_depth, euc_depth_sigs_dataframe, by = "compdepth")

euc_depth_try$compdepth <-factor(euc_depth_try$compdepth,levels=c( "Sp_SurfvsDeep", "Su_SurfvsDeep", "Fa_SurfvsDeep"))

euc_depth_sigs <- euc_depth_try %>%
  select(season1, compdepth, siglabel) %>%
  distinct()

# plot
pdf(file="Figures/Rplot_cDD_euc_depth_boxplot_toptaxa.pdf", width= 5, height=5, pointsize= 14)
euc_depth_boxplot <- ggplot(euc_depth_try, aes(x = compdepth, y = value, fill = compdepth)) + geom_point(size = 2) +
  geom_boxplot() +
  geom_text(data = euc_depth_sigs, aes(label = siglabel, x = compdepth, y = 25), size = 4) +
  scale_fill_manual(name = "", limits=c("Sp_SurfvsDeep", "Su_SurfvsDeep", "Fa_SurfvsDeep"), 
                     values = c("Sp_SurfvsDeep" = "grey","Su_SurfvsDeep" = "grey","Fa_SurfvsDeep" = "grey"))+
  scale_x_discrete(breaks=c("Sp_SurfvsDeep", "Su_SurfvsDeep", "Fa_SurfvsDeep"),
                   labels=c("Spring\n(n=16)", "Summer\n(n=24)", "Fall\n(n=20)")) + 
  xlab("Seasonal distance between cDNA:DNA ratios in surface and deep)") + ylab("Euclidean distance\nlog(cDNA:DNA)") + theme_bw()  +
  theme(plot.title = element_text(face="bold", size = 12),  #Set the plot title
                                 strip.text.x = element_blank(),  #Set the facet titles on x-axis 
                                 strip.text.y = element_blank(),  #Set the facet titles on x-axis 
                                 strip.background = element_blank(),  #Set the facet background to no background
                                 axis.title.x = element_text(face="bold", size=10),  #Set the x-axis title
                                 axis.title.y = element_text(face="bold", size=10),  #Set the y-axis title
                                 axis.text.x = element_text(angle=0, colour = "black", 
                                                            vjust=1, hjust = 1, size=10),  #Set the x-axis labels
                                 axis.text.y = element_text(colour = "black", size=10),  #Set the y-axis labels
                                 legend.title = element_text(size = 8, face="bold"),  #Set the legend title 
                                 legend.text = element_text(size = 8),  #Set the legend text
                                 legend.position="none", #Default the legend position to the right
                                 plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")); euc_depth_boxplot#top, right, bottom, left
dev.off()
```

## top 200 taxa extract OTUs driving patterns PCoA (Fig. 7)
```{r top 200 taxa extract OTUs driving patterns PCoA}
### Aim is to determine those OTUs with similar ratio between Sp PA and FL but diff between Su and Fa PA vs FL
#create phyloseq object with raw data (not log transformed)
cDD_TAX=tax_table(merged_final_toptaxa)
cDD_SAM=sample_data(subset_samples(subset_samples(merged_final_toptaxa,DNA == "cD"),names != "Su.EcD.110.DN"))
cDD_OTU_raw=otu_table(mf_cD2D,taxa_are_rows=TRUE)
cDD_physeq_raw = phyloseq(cDD_OTU_raw,cDD_TAX,cDD_SAM)

#create dataframes to perform stats on
sample_data(cDD_physeq_raw)
cDD_SpPA_noMLB <- data.frame(otu_table(subset_samples(subset_samples(subset_samples(cDD_physeq_raw, Season == "Sp"), Fraction == "E"), Station != "MLB")))
cDD_SpFL_noMLB <- data.frame(otu_table(subset_samples(subset_samples(subset_samples(cDD_physeq_raw, Season == "Sp"), Fraction == "B"), Station != "MLB")))
cDD_SuFaPA_noMLB <- data.frame(otu_table(subset_samples(subset_samples(subset_samples(cDD_physeq_raw, Season != "Sp"), Fraction == "E"), Station != "MLB")))
cDD_SuFaFL_noMLB <- data.frame(otu_table(subset_samples(subset_samples(subset_samples(cDD_physeq_raw, Season != "Sp"), Fraction == "B"), Station != "MLB")))
colnames(cDD_SpPA_noMLB) <- c("SpPA","SpPA","SpPA","SpPA","SpPA","SpPA")
colnames(cDD_SpFL_noMLB) <- c("SpFL","SpFL","SpFL","SpFL","SpFL","SpFL")
colnames(cDD_SuFaPA_noMLB) <- c("SuFaPA","SuFaPA","SuFaPA","SuFaPA","SuFaPA","SuFaPA","SuFaPA","SuFaPA","SuFaPA","SuFaPA","SuFaPA","SuFaPA")
colnames(cDD_SuFaFL_noMLB) <- c("SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL","SuFaFL")
cDD_SpPAFL_noMLB <- cbind(cDD_SpPA_noMLB,cDD_SpFL_noMLB)
cDD_SuFaPAFL_noMLB <- cbind(cDD_SuFaPA_noMLB,cDD_SuFaFL_noMLB)

#run wilcoxon test with Benjamani-Hochberg FDR correction
#cDD_SpPAFL_noMLB_pvals=data.frame(p.adjust(apply(cDD_SpPAFL_noMLB, 1, function(x) wilcox.test(x[1:6],x[7:12])$p.value),"fdr"))
cDD_SpPAFL_noMLB_pvals=data.frame(apply(cDD_SpPAFL_noMLB, 1, function(x) wilcox.test(x[1:6],x[7:12])$p.value))
cDD_SuFaPAFL_noMLB_pvals=data.frame(p.adjust(apply(cDD_SuFaPAFL_noMLB, 1, function(x) wilcox.test(x[1:12],x[13:25])$p.value),"fdr"))

#creat dataframe with log2foldchange mean and sd for each condition
cDD_SpPAFL_noMLB_logmeanPA=data.frame(apply(cDD_SpPAFL_noMLB, 1, function(x) log(mean(x[1:6]),2)))
cDD_SpPAFL_noMLB_logmeanFL=data.frame(apply(cDD_SpPAFL_noMLB, 1, function(x) log(mean(x[7:12]),2)))
cDD_SuFaPAFL_noMLB_logmeanPA=data.frame(apply(cDD_SuFaPAFL_noMLB, 1, function(x) log(mean(x[1:12]),2)))
cDD_SuFaPAFL_noMLB_logmeanFL=data.frame(apply(cDD_SuFaPAFL_noMLB, 1, function(x) log(mean(x[13:25]),2)))

cDD_SpPAFL_noMLB_logsdPA=data.frame(apply(cDD_SpPAFL_noMLB, 1, function(x) log(sd(x[1:6]),2)))
cDD_SpPAFL_noMLB_logsdFL=data.frame(apply(cDD_SpPAFL_noMLB, 1, function(x) log(sd(x[7:12]),2)))
cDD_SuFaPAFL_noMLB_logsdPA=data.frame(apply(cDD_SuFaPAFL_noMLB, 1, function(x) log(sd(x[1:12]),2)))
cDD_SuFaPAFL_noMLB_logsdFL=data.frame(apply(cDD_SuFaPAFL_noMLB, 1, function(x) log(sd(x[13:25]),2)))

cDD_wilcox <- cbind(cDD_SpPAFL_noMLB_logmeanPA,cDD_SpPAFL_noMLB_logmeanFL,cDD_SuFaPAFL_noMLB_logmeanPA,cDD_SuFaPAFL_noMLB_logmeanFL,cDD_SpPAFL_noMLB_logsdPA,cDD_SpPAFL_noMLB_logsdFL,cDD_SuFaPAFL_noMLB_logsdPA,cDD_SuFaPAFL_noMLB_logsdFL,cDD_SpPAFL_noMLB_pvals,cDD_SuFaPAFL_noMLB_pvals)
colnames(cDD_wilcox) <- c("SpPA_mean","SpFL_mean","SuFaPA_mean","SuFaFL_mean","SpPA_sd","SpFL_sd","SuFaPA_sd","SuFaFL_sd","SpPAFL_pvals","SuFaPAFL_pvals")
cDD_wilcox$OTU <- rownames(cDD_wilcox)

#set NaN as p value of 1 and filter p-values to identify thsoe not diff between Sp PA and FL but diff between SuFa PA and FL
cDD_wilcox$SpPAFL_pvals[cDD_wilcox$SpPAFL_pvals=="NaN"]<-1.00
cDD_wilcox$SuFaPAFL_pvals[cDD_wilcox$SuFaPAFL_pvals=="NaN"]<-1.00
cDD_wilcox_simSpPAFL <- filter(cDD_wilcox, as.numeric(SpPAFL_pvals) > 0.10)
cDD_wilcox_simSpPAFL_diffSuFaPAFL <- filter(cDD_wilcox_simSpPAFL, as.numeric(SuFaPAFL_pvals) < 0.01)
rownames(cDD_wilcox_simSpPAFL_diffSuFaPAFL) <- cDD_wilcox_simSpPAFL_diffSuFaPAFL$OTU
#View(cDD_wilcox_simSpPAFL_diffSuFaPAFL)

#create physeq object with this subset of otus
cDD_physeq_diffeat <- subset_taxa(cDD_physeq, Species %in% rownames(cDD_wilcox_simSpPAFL_diffSuFaPAFL))
cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax <- cbind(cDD_wilcox_simSpPAFL_diffSuFaPAFL,tax_table(cDD_physeq_diffeat))

#generate dataframe with taxa info and log2foldchange + column describing which condition (Habitat) this log2foldchange stems from
t1 <- data.frame(c(cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax[1],cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax[13:19]))
t1$Habitat <- "SpPA"
colnames(t1)[1] <- "log2FoldChange"
rownames(t1) <- t1$Species
t2 <- data.frame(c(cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax[2],cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax[13:19]))
t2$Habitat <- "SpFL"
colnames(t2)[1] <- "log2FoldChange"
rownames(t2) <- t2$Species
t3 <- data.frame(c(cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax[3],cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax[13:19]))
t3$Habitat <- "SuFaPA"
colnames(t3)[1] <- "log2FoldChange"
rownames(t3) <- t3$Species
t4 <- data.frame(c(cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax[4],cDD_wilcox_simSpPAFL_diffSuFaPAFL_wtax[13:19]))
t4$Habitat <- "SuFaFL"
colnames(t4)[1] <- "log2FoldChange"
rownames(t4) <- t4$Species
t124 <- rbind(t1,t2,t4)
t124avg <- ddply(t124, ~Species, function(x){data.frame(t124_mean = mean(as.numeric(x$log2FoldChange)))})
t124sd <- ddply(t124, ~Species, function(x){data.frame(t124_sd = sd(as.numeric(x$log2FoldChange)))})
t124summ <- cbind(t1,t124avg$t124_mean,t124sd$t124_sd) %>% select(10,2,3,4,5,6,7,8,9,11)
t124summ$Habitat <- "SpFLSpPASuFaFL"
colnames(t124summ)[1] <- "log2FoldChange"
colnames(t124summ)[10] <- "stdev"
t3$stdev = 0
#cDD_diff <- rbind(t1,t2,t3,t4)
cDD_diff <- rbind(t124summ,t3)

#now plot data, sorting the otus based on how different SuFaPA ratio is from SpPA --- FOR FIGURE 7
y = tapply(abs(t3$log2FoldChange-t1$log2FoldChange)-abs(t2$log2FoldChange-t1$log2FoldChange), t3$Species, function(x) max(x))
y = sort(y, TRUE)
cDD_diff$Species = factor(as.character(cDD_diff$Species), levels=names(y))
pdf(file="Figures/Rplot_cDD_PCoAdrivingOTUsSpvsSuFa_noMLB.pdf", width= 10, height=7.5, pointsize= 14, useDingbats=FALSE)
drivingOTUsSpvsSuFa_noMLB <- ggplot(cDD_diff, aes(x=Species, y=log2FoldChange, color=Habitat, shape = Habitat)) + 
    geom_point(alpha=1, size=4) + theme_bw() +  #ggtitle(title) +  
    scale_color_manual(name = "Habitat", breaks=c("SpFLSpPASuFaFL",  "SuFaPA"),
                     labels = c("Mean(SpFL,SpPA,SuFaFL)",  "SuFaPA"), 
                     values = c("SpFLSpPASuFaFL" = "black", "SuFaPA" = "darkorange")) +
    scale_shape_manual(name = "Habitat", breaks=c("SpFLSpPASuFaFL",  "SuFaPA"),
                     labels = c("Mean(SpFL,SpPA,SuFaFL)",  "SuFaPA"), 
                     values = c("SpFLSpPASuFaFL" = 16,  "SuFaPA" = 17)) +
    geom_errorbar(aes(ymin=log2FoldChange-stdev, ymax=log2FoldChange+stdev), width=.5) +
    scale_y_continuous(breaks=seq(-3,3,1),limits=c(-3,3)) + 
    theme(axis.title.x = element_text(face="bold", size=16),
          axis.text.x = element_text(angle=30, colour = "black", vjust=1, hjust = 1, size=12),
          axis.text.y = element_text(colour = "black", size=16),
          axis.title.y = element_text(face="bold", size=16),
          plot.title = element_text(face="bold", size = 20),
          legend.title = element_text(size=12, face="bold"),
          legend.text = element_text(size = 12),
          strip.background = element_rect(colour="black"),
          legend.position="right"); drivingOTUsSpvsSuFa_noMLB
dev.off()

#output sorted otus and taxonomy for table generation
View(levels(cDD_diff$Species))
View(data.frame(cDD_diff$Species,cDD_diff$Phylum,cDD_diff$Class,cDD_diff$Order,cDD_diff$Family,cDD_diff$Genus,cDD_diff$Tribe))
```

## top 200 taxa identify significantly more active than present OTUs (Fig. 4, Table 3)
```{r identify significantly more active than present OTUs}
########## cDNA vs DNA DEseq ##########

phylum.colors<-group.colors

# deseq without pruning low abundance - otu level - per fraction - PA - no scaling (done by DEseq2!)
raw_merged_PA <- subset_samples(raw_merged_toptaxa, Fraction != "B")
deseq_otu_DNAcDNA_PA_raw <- deSEQ_noprune(raw_merged_PA, ~ DNA)
deseq_otu_DNAcDNA_PA_raw$plotvalue <- as.vector(as.numeric(taxa_sums(raw_merged_PA)/sum(sample_sums(raw_merged_PA))*100)^.25) # abundance scaled to optimize gradient of glyph sizes in image
deseq_otu_DNAcDNA_PA_raw$sig <- as.factor((deseq_otu_DNAcDNA_PA_raw$padj<0.01)*1)
deseq_otu_DNAcDNA_plot_PA_raw <- plot_deSEQ(deseq_otu_DNAcDNA_PA_raw, "OTU cDNA:DNA PA samples")
deseq_otu_DNAcDNA_plot_PA_raw

# deseq without pruning low abundance - otu level - per fraction - FL - no scaling (done by DEseq2!)
raw_merged_FL <- subset_samples(raw_merged_toptaxa, Fraction != "E")
deseq_otu_DNAcDNA_FL_raw <- deSEQ_noprune(raw_merged_FL, ~ DNA)
deseq_otu_DNAcDNA_FL_raw$plotvalue <- as.vector(as.numeric(taxa_sums(raw_merged_FL)/sum(sample_sums(raw_merged_FL))*100)^.25)
deseq_otu_DNAcDNA_FL_raw$sig <- as.factor((deseq_otu_DNAcDNA_FL_raw$padj<0.01)*1)
deseq_otu_DNAcDNA_plot_FL_raw <- plot_deSEQ(deseq_otu_DNAcDNA_FL_raw, "OTU cDNA:DNA FL samples")
deseq_otu_DNAcDNA_plot_FL_raw

#plot together --- FOR FIGURE 4
Phylum_temp <- as.character(deseq_otu_DNAcDNA_PA_raw$Phylum)
for  (i in 1:length(Phylum_temp)){ 
  Phylum_temp[i] = paste(Phylum_temp[i],"PA",sep=" ")
} 
deseq_otu_DNAcDNA_PA_raw$Phylum_plot <- Phylum_temp

Phylum_temp <- as.character(deseq_otu_DNAcDNA_FL_raw$Phylum)
for  (i in 1:length(Phylum_temp)){ 
  Phylum_temp[i] = paste(Phylum_temp[i],"FL",sep=" ")
} 
deseq_otu_DNAcDNA_FL_raw$Phylum_plot <- Phylum_temp

deseq_otu_DNAcDNA_FLandPA_raw<-rbind(deseq_otu_DNAcDNA_FL_raw,deseq_otu_DNAcDNA_PA_raw)

pdf(file="Figures/Rplot_cDvsD_FL_noscale_top200.pdf", width= 10, height=15, pointsize= 14, useDingbats=FALSE)
deseq_otu_DNAcDNA_FLandPA_raw$Phylum_plot <-factor(deseq_otu_DNAcDNA_FLandPA_raw$Phylum_plot,levels=c("Proteobacteria FL","Proteobacteria PA","Alphaproteobacteria FL","Alphaproteobacteria PA","Betaproteobacteria FL","Betaproteobacteria PA","Gammaproteobacteria FL","Gammaproteobacteria PA","Deltaproteobacteria FL","Deltaproteobacteria PA","Bacteroidetes FL","Bacteroidetes PA","Cytophagia FL","Cytophagia PA","Flavobacteria FL","Flavobacteria PA","Sphingobacteria FL","Sphingobacteria PA","Actinobacteria FL","Actinobacteria PA","Cyanobacteria FL","Cyanobacteria PA","Verrucomicrobia FL","Verrucomicrobia PA","Planctomycetes FL","Planctomycetes PA","Chloroflexi FL","Chloroflexi PA","Acidobacteria FL","Acidobacteria PA","Armatimonadetes FL","Armatimonadetes PA","Chlorobi FL","Chlorobi PA","Firmicutes FL","Firmicutes PA","Gemmatimonadetes FL","Gemmatimonadetes PA","unclassified FL","unclassified PA"))
deseq_otu_DNAcDNA_plot_FLandPA_raw <- plot_deSEQ_combo(deseq_otu_DNAcDNA_FLandPA_raw, "OTU cDNA:DNA FL and PA samples")
deseq_otu_DNAcDNA_plot_FLandPA_raw
dev.off()

####pull out most different, highly active OTUs  --- FOR TABLE 3
View(deseq_otu_DNAcDNA_FLandPA_raw)

```

## top 200 taxa identify significantly more active than present OTUs - No Actinobacteria (to determine effect of Actinobacteria, no figure in paper, but mentioned in results)
```{r identify significantly more active than present OTUs - No Actinobacteria}
### cDNA vs DNA DEseq

#repeat the DEseq analysis without Actinobacteria
raw_merged_toptaxa_noAct <- subset_taxa(raw_merged_toptaxa, Phylum != "Actinobacteria")

phylum.colors<-group.colors

# deseq without pruning low abundance - otu level - per fraction - PA - no scaling (done by DEseq2!)
raw_merged_PA <- subset_samples(raw_merged_toptaxa_noAct, Fraction != "B")
deseq_otu_DNAcDNA_PA_raw <- deSEQ_noprune(raw_merged_PA, ~ DNA)
deseq_otu_DNAcDNA_PA_raw$plotvalue <- as.vector(as.numeric(taxa_sums(raw_merged_PA)/sum(sample_sums(raw_merged_PA))*100)^.25)
deseq_otu_DNAcDNA_PA_raw$sig <- as.factor((deseq_otu_DNAcDNA_PA_raw$padj<0.01)*1)
deseq_otu_DNAcDNA_plot_PA_raw <- plot_deSEQ(deseq_otu_DNAcDNA_PA_raw, "OTU cDNA:DNA PA samples")
deseq_otu_DNAcDNA_plot_PA_raw

# deseq without pruning low abundance - otu level - per fraction - FL - no scaling (done by DEseq2!)
raw_merged_FL <- subset_samples(raw_merged_toptaxa_noAct, Fraction != "E")
deseq_otu_DNAcDNA_FL_raw <- deSEQ_noprune(raw_merged_FL, ~ DNA)
deseq_otu_DNAcDNA_FL_raw$plotvalue <- as.vector(as.numeric(taxa_sums(raw_merged_FL)/sum(sample_sums(raw_merged_FL))*100)^.25)
deseq_otu_DNAcDNA_FL_raw$sig <- as.factor((deseq_otu_DNAcDNA_FL_raw$padj<0.01)*1)
deseq_otu_DNAcDNA_plot_FL_raw <- plot_deSEQ(deseq_otu_DNAcDNA_FL_raw, "OTU cDNA:DNA FL samples")
deseq_otu_DNAcDNA_plot_FL_raw

#plot together
Phylum_temp <- as.character(deseq_otu_DNAcDNA_PA_raw$Phylum)
for  (i in 1:length(Phylum_temp)){ 
  Phylum_temp[i] = paste(Phylum_temp[i],"PA",sep=" ")
} 
deseq_otu_DNAcDNA_PA_raw$Phylum_plot <- Phylum_temp

Phylum_temp <- as.character(deseq_otu_DNAcDNA_FL_raw$Phylum)
for  (i in 1:length(Phylum_temp)){ 
  Phylum_temp[i] = paste(Phylum_temp[i],"FL",sep=" ")
} 
deseq_otu_DNAcDNA_FL_raw$Phylum_plot <- Phylum_temp

deseq_otu_DNAcDNA_FLandPA_raw<-rbind(deseq_otu_DNAcDNA_FL_raw,deseq_otu_DNAcDNA_PA_raw)

pdf(file="Figures/Rplot_cDvsD_FL_noscale_top200_noAct.pdf", width= 10, height=7.5, pointsize= 14, useDingbats=FALSE)
deseq_otu_DNAcDNA_FLandPA_raw$Phylum_plot <-factor(deseq_otu_DNAcDNA_FLandPA_raw$Phylum_plot,levels=c("Proteobacteria FL","Proteobacteria PA","Alphaproteobacteria FL","Alphaproteobacteria PA","Betaproteobacteria FL","Betaproteobacteria PA","Gammaproteobacteria FL","Gammaproteobacteria PA","Deltaproteobacteria FL","Deltaproteobacteria PA","Bacteroidetes FL","Bacteroidetes PA","Cytophagia FL","Cytophagia PA","Flavobacteria FL","Flavobacteria PA","Sphingobacteria FL","Sphingobacteria PA","Actinobacteria FL","Actinobacteria PA","Cyanobacteria FL","Cyanobacteria PA","Verrucomicrobia FL","Verrucomicrobia PA","Planctomycetes FL","Planctomycetes PA","Chloroflexi FL","Chloroflexi PA","Acidobacteria FL","Acidobacteria PA","Armatimonadetes FL","Armatimonadetes PA","Chlorobi FL","Chlorobi PA","Firmicutes FL","Firmicutes PA","Gemmatimonadetes FL","Gemmatimonadetes PA","unclassified FL","unclassified PA"))
deseq_otu_DNAcDNA_plot_FLandPA_raw <- plot_deSEQ_combo(deseq_otu_DNAcDNA_FLandPA_raw, "OTU cDNA:DNA FL and PA samples")
deseq_otu_DNAcDNA_plot_FLandPA_raw
dev.off()
```


